---
title: "DESEQ_R_PE_HDAC"
author: "Natalia Garcia Sanchez"
date: "2023-06-05"
output: html_document
---

```{r Load Packages, include=FALSE}
library("DESeq2")
library("apeglm")
library("pheatmap")
library("tidyverse")
require("vidger")
library("tidyverse")
library("biomaRt")
library("ggplot2")
library("EnhancedVolcano")
library("RColorBrewer")
library("gplots")
library("amap")
library("ggplot2")
library("BiocParallel")

suppressWarnings(suppressMessages({
  library('easypackages',warn.conflicts = FALSE, quietly=TRUE)
  library("readr",warn.conflicts = FALSE, quietly=TRUE)
  library("ggplot2",warn.conflicts = FALSE, quietly=TRUE)
  library("RColorBrewer",warn.conflicts = FALSE, quietly=TRUE)
  library("biomaRt",warn.conflicts = FALSE, quietly=TRUE)
  library("DESeq2",warn.conflicts = FALSE, quietly=TRUE)
  library("pheatmap",warn.conflicts = FALSE, quietly=TRUE)
  library("RColorBrewer",warn.conflicts = FALSE, quietly=TRUE)
  library("viridis",warn.conflicts = FALSE, quietly=TRUE)
  library("ggpubr",warn.conflicts = FALSE, quietly=TRUE)
  library('EnhancedVolcano',warn.conflicts = FALSE, quietly=TRUE)}))

```

# Differential Expression analysis description

Differential Gene Expression (DGE) analysis will consist of simply two
lines of code:

1.  The first will call the DESeq function on a DESeqDataSet object that
    we've just created under the name dds. It will be returned under the
    same R object name dds.
2.  The results are extracted using the results function on the dds
    object and results will be extracted as a table under the name res
    (short for results).

# Methods

*Pairwise Differential Expression with Counts in R using DESeq2*

Statistical analysis with `DESeq()` function will perform the following
steps of the DGE pipeline:

-   **Estimating size factors + Normalization** : This method of
    normalization consists on calculating a *scaling size factor* based
    on the *mean of the ratios* of each expression level per sample and
    gene with the Geometric Mean of expression levels across genes in
    the dataset. This part of the analysis accounts for the fact that
    standard RNA-seq measures relative abundances of transcripts within
    each sample, not absolute abundances (i.e. transcripts/cell).
    Because of this, if libraries are sequenced to different depths,
    genes with identical expression levels will have different counts.
    It may seem that simply adjusting for sequencing depth could account
    for this issue, but changes in gene expression among samples
    (library composition) account for another level of sequencing bias,
    so a slightly more complex normalization is applied.
-   **Estimating dispersion through *three dispersion estimate steps***:
    This is necessary in a High Throughput Sequence (HTS) experiment, as
    genes with lower levels of counts will also present more expression
    variability, and present difficulties for fitting a Generalized
    Linear Model accurately. DESeq2 models the variance in the counts
    for each sample using a Negative Binomial distribution. In this
    distribution the variance in counts is determined by a dispersion
    parameter. This parameter needs to be estimated for each gene, but
    for sample sizes typical to RNA-Seq (3-5 per treatment) there isn't
    a lot of power to estimate it accurately. The key innovation of
    packages like DESeq2 is to assume genes with similar expression have
    similar dispersions, and to pool information across them to estimate
    the dispersion for each gene. These steps accomplish that using an
    "empirical Bayesian" Maximum Likelihood Estimation (MLE) procedure.
    This establishes that the dispersion parameter for a particular gene
    is a compromise between the signal in the data for that gene, and
    the signal in other, similar genes.
-   **Fitting a Generalized Linear Model** and **Statistical hypothesis
    testing**: Once the dispersions are estimated and the size factors
    are accounted for fitting in a Generalized Linear Model from which
    log2 fold change (LFC) values are extrapolated, this part of the
    analysis asks if there are treatment effects for each gene. In the
    default case, DESeq2 uses a Wald parametric test to ask if the LFC
    between two treatments is significantly different than zero. This
    test is used upon performing hypothesis testing a the set of
    predictors used in to a variety of models on binary or continuous
    variables, as an approximation of the Likelihood Ratio Test.

We will also control statistical inference results through **False
Discovery Rate (FDR) control**. This step adjust the previously
calculated p-values in the hypothesis testing as a control of type I
errors (False Positive Errors occurring as a result of the latter
hypothesis testing across multiple comparisons). Along these lines,
Benjamini-Hochberg: Benjamini and Hochberg (BH) (1995) defined the
concept of FDR and created an algorithm to control the expected FDR
below a specified level given a list of independent p-values. An
interpretation of the BH method for controlling the FDR is implemented
in DESeq2 in which we rank the genes by p-value, then multiply each
ranked p-value by m/rank.

# Results

The pipeline will return csv results of normalized counts, their mean
across samples (baseMean), LFC values (log2FoldChange) and a Standard
Error of this measure in the GLM (lfcSE), the Wald statistic value
(stat) and p-values(pvalue) from hypothesis testing, and the p-adj (FDR
q-value) from the BH FDR control method (padj), complete and filtered by
set of genes with significant differential expression (pvalue\<0.05).

Once the statistical inference is done, a preliminary exploration of the
Differentially Expressed Genes is performed. In this way, the present
script will also undertake exploratory data analysis of RNAseq data with
DESeq2 for a variety of visualization, QC and other plots to get a sense
of what the RNAseq data looks like based on DESEq2 analysis.

1)  Expression of gene across the samples- what does expression look
    like across genes? Histogram of the log scaled counts and filtering
    the low expressed genes
2)  Clustering analysis (PCA, Hierarchical clustering, Heatmap)
3)  Significant Genes - MA plot
4)  Significant Genes - Volcano plot

------------------------------------------------------------------------

# Loading data

Loading data for HiSAT2 alignment standarization and Differential
Expression analysis.

Listing files from input directory, accounting for the order of the
files. Note that the detected counts have the `*.counts`

```{r read count files}
# Create an object with the directory containing HTseq counts:
directory <- "/home/famgarcia/TFM/data/Testillano col/MV vs PE_HDAC/counts"

# list.files(directory) will return the files in the input directory
sampleFiles <- list.files(directory, pattern = ".*counts")
sampleFiles
```

Reading merged counts for standarization (mean of ratios) pipeline,
accounting for the order of the files.

First it is required to create a sample table accounting for the sample
names, file names, experimental conditions they face, and biological
replicate number

```{r create vector samplefiles}

# create a vector of sample names, ensuring that these are in the same order as the "sampleFiles" object.
sampleNames <- c("Trep1", "Trep2", "Trep3", "VMrep1", "VMrep2", "VMrep3")

# create a vector of conditions in order, ensuring that these are in the same order as the "sampleFiles" object.
replicate <- c("Rep1","Rep2", "Rep3","Rep1","Rep2", "Rep3")

# create a vector of conditions. again, mind that they are ordered correctly!
sampleCondition <- c("T","T", "T","VM", "VM", "VM")

# Passing informative condition, replicate, sample file, and sample name vectors onto a sampleTable ready for reading by DeSeq2
sampleTable <- data.frame(
		sampleName = sampleNames,
		fileName = sampleFiles,
		condition = sampleCondition,
		replicate = replicate)

```

## Make DESeq2 object from counts and metadata

Creating DESeqDataSet object with directory passing the metadata from
sampleTable and the count directory files

Create an object with the directory containing HTseq counts, specifying
the control condition `C`:
```{r}


countTable <- read.csv("/home/famgarcia/VM_vs_HDAC_DEA/count_matrix.csv")
colnames(countTable)<-c("gene_id","VMrep1","VMrep2","VMrep3","Trep1","Trep2","Trep3")
rownames(countTable) <-countTable[,1]
countTable$gene_id <- NULL

condition <- data.frame(
  condition<-c("VM","VM","VM","T","T","T"))
rownames(condition) <-c("VMrep1","VMrep2","VMrep3","Trep1","Trep2","Trep3")
ncol(countTable) == nrow(condition)
  
  
dds<-DESeqDataSetFromMatrix(
    countData=as.matrix(countTable),
    design =~ condition,
    colData=data.frame(
      condition = c("VM","VM","VM","T","T","T")))
```

```{r}

#specifying the reference level:

dds$condition <- factor(dds$condition, levels = c("VM", "T"))
dds$condition

dds$condition <- relevel(dds$condition, ref = "VM")
dds$condition

```

## Pre-filtering

Here we perform a minimal pre-filtering to keep only rows that have at
least 10 reads total. This will give an insight of what expression looks
like across genes

```{r}

# sum counts for each gene across samples
sumcounts <- rowSums(counts(dds))

# take the log
logsumcounts <- log(sumcounts,base=10)
sprintf("Genes before filtering : %s",length(logsumcounts))

# plot a histogram of the log scaled counts
plot_without_filter <- hist(logsumcounts,breaks=100, main="Histogram of the log scaled counts")

# get genes with summed counts greater than 10; remove lowly expressed genes
keep <- sumcounts > 10

# keep only the genes for which the vector "keep" is TRUE
dds_filter <- dds[keep,]

sumcounts <- rowSums(counts(dds_filter))
# take the log
logsumcounts <- log(sumcounts,base=10)
# plot a histogram of the log scaled counts


plot_with_filter <- hist(logsumcounts,breaks=100,main="Histogram of the log scaled counts \n keep only the genes with summed counts greater than 10" )
```
```{r}
write.table(as.data.frame(rownames(dds_filter)),"/home/famgarcia/VM_vs_HDAC_DEA/filtered_VM_vs_T.csv", row.names=FALSE, quote=FALSE, col.names=FALSE)
```

## Run DESeq2 DEG analysis

By running DESeq over the filtered HTSeqCount DESeq2 object, we will:

-   **Normalize the count expression data** with the mean of ratios
    method. This method accounts for biases in sequencing depth and gene
    composition.

-   **Estimate dispersion levels per gene**. ( Negative Binomial
    distribution *see description of how DESeq2 performs this step
    above* ).

-   **Fit a Generalized Linear Model** to Log2FoldChange \~
    log2(treatment/baseline expression) and perform hypothesis testing
    of differential expression

Create the DESeq data object:

PlusP samples from each type of cell group is used as a reference level
(control)

```{r}

# normalize and calculate dispersion
dds <- DESeq(dds_filter) 

plotDispEsts(dds, main="Dispersion plot")

```

# Get basic statisics about the number of significant genes

A positive log2 fold change for a gene would mean that this gene is more
abundant in treatment condition than in the control condition.

Extracting the table of differential genes and Exploratory data analysis
of RNAseq data with DESeq2 QC plots

-   A log2 equal to 1 means that gene X has a higher expression (x2,
    two-fold) in the treatment condition compared to the control
    condition.

-   A log2 equal to -1 means that gene X has a smaller expression in the
    treatment condition.

-   Example: untreated (control) = 230 treated= 750
    log2(treated/untreated) \# equals 1.705257

**Column Result Summary**

baseMean: mean of normalized counts for all samples log2FoldChange: log2
fold change lfcSE: standard error stat: Wald statistic pvalue: Wald test
p-value padj: BH adjusted p-values multiple testing

```{r}


#Get basic statisics about the number of significant genes

res <- results(dds, pAdjustMethod="BH") 

resultsNames(dds)

summary(res)

# check out the first few lines

head(res)

mcols(res, use.names = T)


```
```{r}
ressign<-subset(res,abs(log2FoldChange)>1)
sprintf('genes with padj<0.05 and |LFC|>1 : %s', nrow(ressign[which((ressign$padj < 0.05) & (!is.na(ressign$padj))),]))
sprintf('genes with padj<0.05 and LFC>1 (up) : %s', nrow(ressign[which((ressign$log2FoldChange>1 & ressign$padj < 0.05) & (!is.na(ressign$padj))),]))
sprintf('genes with padj<0.05 and LFC<-1 (down) : %s', nrow(ressign[which((!ressign$log2FoldChange>1) & ressign$padj < 0.05 & (!is.na(ressign$padj))),]))
```

## CREATE OUTPUTS

```{r}

##Create normalized read counts
normalized_counts <- counts(dds, normalized=TRUE)

normalized_counts_mad <- apply(normalized_counts, 1, mad)
normalized_counts <- normalized_counts[order(normalized_counts_mad, decreasing=T), ]


#DESeq get results table - TOTAL RESULTS


Res_A_X_total <- results(dds,
name="condition_T_vs_VM",
pAdjustMethod="BH")

Res_A_X_total <- Res_A_X_total[order(Res_A_X_total$padj),]


#########################
# Query the mart/dataset
#########################

# see a list of "marts" available at host "plants.ensembl.org"
listMarts <-listMarts(host="https://plants.ensembl.org")

# create an object 
mart <- useMart(biomart="plants_mart", host="https://plants.ensembl.org")

# see a list of data sets within the mart

listdatasets <- listDatasets(mart)
(listDatasets(mart))[grep("bnapus",listDatasets(mart)[,1]),]

# create an object for the plants_mart-Ensembl 

bnapus_mart <- useMart(biomart = "plants_mart", host = "https://plants.ensembl.org", dataset = "bnapus_eg_gene"	)


# see a list of all "attributes" available

attributes <- listAttributes(mart = bnapus_mart, page="feature_page")
attributes
```

```{r}


# get gene names and transcript lengths when they exist
ann <- getBM(filter="ensembl_gene_id",value=rownames(Res_A_X_total),attributes=c("ensembl_gene_id","chromosome_name", "external_gene_name","start_position", "end_position", "strand", "transcript_biotype","transcript_length", "description"),mart=bnapus_mart, useCache = FALSE)

# pick only the longest transcript for each gene ID
ann <- group_by(ann, ensembl_gene_id) %>% 
  summarize(.,description=unique(description),transcript_length=max(transcript_length))

# Reannotate transcripts
go_ann <- getBM(filter="ensembl_gene_id",value=rownames(Res_A_X_total),attributes=c("ensembl_gene_id","chromosome_name", "external_gene_name","start_position", "end_position", "strand", "transcript_biotype","transcript_length", "description"),bnapus_mart, useCache = FALSE)


nrow(Res_A_X_total)
nrow(ann)


Res_A_X_total <- data.frame(Res_A_X_total)
Res_A_X_total  <- rownames_to_column(Res_A_X_total, var = "ensembl_gene_id")


#nrow(ann)
#nrow(Res_A_X_total)

Res_A_X_total <-merge(Res_A_X_total,go_ann,by='ensembl_gene_id',all.x = TRUE, all.y = TRUE)

#Export output files
Res_A_X_total$sig <- ifelse(Res_A_X_total$padj <= 0.05, "yes", "no")

Res_A_X_total_0.05 <-subset(Res_A_X_total, padj <=0.05)



#dir.create(paste(directory,"/DE_results/",sep=""))
write.table(normalized_counts,"/home/famgarcia/TFM/data/Testillano col/MV vs PE_HDAC/Condition_T_vs_VM_normcounts.csv",quote=FALSE, sep=",")


#Note on p-values set to NA: some values in the results table can be set to NA for one of the following reasons:

#If within a row, all samples have zero counts, the baseMean column will be zero, and the log2 fold change estimates, p value and adjusted p value will all be set to NA.
#If a row contains a sample with an extreme count outlier then the p value and adjusted p value will be set to NA. These outlier counts are detected by Cook’s distance. Customization of this outlier filtering and description of functionality for replacement of outlier counts and refitting is described below
#If a row is filtered by automatic independent filtering, for having a low mean normalized count, then only the adjusted p value will be set to NA. Description and customization of independent filtering is described below
```


```{r}
n_occur <- data.frame(table(Res_A_X_total$external_gene_name))
Res_duplicated_GN <- Res_A_X_total[Res_A_X_total$external_gene_name %in% n_occur$Var1[n_occur$Freq > 1],]
#sort dataframe
Res_duplicated_GN<-Res_duplicated_GN[order(Res_duplicated_GN$external_gene_name),]
listch <- Res_duplicated_GN$external_gene_name

i<-0
k<-0
newlistch<-vector()
for (gene in listch){
  k=k+1
  if(k==1){lastgene<-gene}
  else{lastgene<-listch[k-1]}
  
  if (gene == lastgene){
    i<-i+1
    gene<-paste(gene, i, sep="_")
    newlistch <- c(newlistch, gene)
  }else{
    i=1
    gene<-paste(gene, i, sep="_")
    newlistch <- c(newlistch, gene)
    }
}

Res_duplicated_GN$external_gene_name <- newlistch

Res_non_duplicated_GN <- Res_A_X_total[!Res_A_X_total$external_gene_name %in% n_occur$Var1[n_occur$Freq > 1],]

Res_A_X_total_plot =rbind(Res_non_duplicated_GN, Res_duplicated_GN)


```
```{r}
# Save file with sorted ranking score - Wald statistic for ranking
rank_res <- Res_A_X_total_plot[,c(9,5)]
write.table(rank_res[order(rank_res$stat,decreasing=TRUE),],"/home/famgarcia/TFM/data/Testillano col/MV vs PE_HDAC/Rank_T_vs_VM.rnk", row.names = FALSE, sep=",", quote=FALSE, col.names=FALSE)
```

```{r}
rownames(Res_A_X_total_plot)  <- Res_A_X_total_plot$external_gene_name
```

```{r}
head(Res_A_X_total_plot)
```

## RNA-seq data quality assessment

### Sample Clustering {.tabset .tabset-fade .tabset-pills}

Hierarchical clustering RNA-seq gene expression patterns based on sample
distance (Euclidean distance matrix) of log transformed expression from
transcripts based on DeSeq2 (Love et al. 2014).

```{r echo=, message=FALSE, warning=FALSE}
# log transformed expression

#This function transforms the count data to the log2 scale in a way which minimizes differences between samples for rows with small counts, and which normalizes with respect to library size. The rlog transformation produces a similar variance stabilizing effect as varianceStabilizingTransformation, though rlog is more robust in the case when the size factors vary widely. The transformation is useful when checking for outliers or as input for machine learning techniques such as clustering or linear discriminant analysis. rlog takes as input a DESeqDataSet and returns a RangedSummarizedExperiment object.

rld <- rlog(dds, blind=FALSE)
rlogMat <- assay(rld)
rlogMat <- rlogMat[order(normalized_counts_mad, decreasing=T), ]

# Generar color
hmcol <- colorRampPalette(brewer.pal(9, "GnBu"))(100)

 # Pearson correlation
pearson_cor <- as.matrix(cor(rlogMat, method="pearson"))

# Hierarchical clustering
hc <- hcluster(t(rlogMat), method="pearson")

svg("/home/famgarcia/TFM/data/Testillano col/MV vs PE_HDAC/Treecorr_ME_HDAC.svg")
TreeC = as.dendrogram(hc, method="average")
plot(TreeC,
     main = "Sample Clustering-checking outliers ",
     ylab = "Height")
dev.off()

#adjust plot margins
#par(mar=c(7,4,9,2)+0.1) 

svg("/home/famgarcia/TFM/data/Testillano col/MV vs PE_HDAC/HeatmapRLDcorr_ME_HDAC.svg")
heatmap.2(pearson_cor, Rowv=as.dendrogram(hc), symm=T,trace="none",
col=hmcol, main="Pearson correlation of each
sample")
dev.off()
```
```{r}
write.csv(pearson_cor, "/home/famgarcia/TFM/data/Testillano col/MV vs PE_HDAC/pearsonRLD_ME_HDAC.csv")
```

```{r}

```


```{r echo=, message=FALSE, warning=FALSE}

# normalized, variance-stabilized transformed counts for visualization
vsd <- vst(dds, blind=FALSE)

svg("/home/famgarcia/TFM/data/Testillano col/MV vs PE_HDAC/VST_PCA_ME_HDAC.svg")
plotPCA(vsd, intgroup="condition") + geom_text(aes(label=name),vjust=-1) +  ggtitle("PCA Plot using variance stabilizing transformation data") + scale_x_continuous(limits = c(-100, 100))  +  scale_y_continuous(limits = c(-50, 50)) # change limits to fix figure dimensions
dev.off()

svg("/home/famgarcia/TFM/data/Testillano col/MV vs PE_HDAC/RLD_PCA_ME_HDAC.svg")
plotPCA(rld, intgroup="condition") + geom_text(aes(label=name),vjust=1.5) + ggtitle("PCA Plot using Regularized log  Transformation data") + scale_x_continuous(limits = c(-200, 200))  +  scale_y_continuous(limits = c(-100, 100)) # change limits to fix figure dimensions
dev.off()

pca_data <- plotPCA(rld, intgroup="condition", returnData=T)
```

```{r}
listddsgenes <- rownames(dds)
dds_new <- dds
rownames(dds_new) <- rownames(Res_A_X_total_plot[match(listddsgenes,Res_A_X_total_plot$ensembl_gene_id),])

```

### PlotMA {.tabset .tabset-fade .tabset-pills}

In DESeq2, the function plotMA shows the log2 fold changes attributable
to a given variable over the mean of normalized counts for all the
samples in the DESeqDataSet. Points will be colored red if the adjusted
p value is less than 0.05. Points which fall out of the window are
plotted as open triangles pointing either up or down.

A standard Bland-Altman MA plot and a MA plot with shrunken LFC values
will be plotted. In this latter plot, shrinkage of effect size (LFC
estimates) is useful for visualization and ranking of genes. It is more
useful visualize the MA-plot for the shrunken log2 fold changes, which
remove the noise associated with log2 fold changes from low count genes
without requiring arbitrary filtering thresholds. This helps to get more
meaningful log2 fold changes for all genes independently of their
expression level.

```{r}
# This plot gives an idea of the dispersion variability among genes that have high gene counts versus genes that do not. This will affect the LFC, which is why it is important to calculate a prior distribution with LFC that are expected not to change(for genes with high counts), and use it to shrink or predict new estimates of LFC values with reduced noise for genes with less counts.
plotSparsity(dds_new)
```

```{r, warning = FALSE, message = FALSE, echo = FALSE, fig.width=5, fig.height=5}
svg("/home/famgarcia/TFM/data/Testillano col/MV vs PE_HDAC/PMA_Standard_PE_HDAC.svg")
plotMA(dds_new , alpha =0.05, main=("Standard Bland-Altman- MA Plot"))
dev.off()

resLFC <- lfcShrink(dds_new, coef="condition_T_vs_VM", type="apeglm")

svg("/home/famgarcia/TFM/data/Testillano col/MV vs PE_HDAC/PMA_ShrunkenME_HDAC.svg")
#You can see that genes with low counts are now shrinked.                  
plotMA(resLFC, alpha =0.05, ylim=c(-4,4) ,main= ("MA plot shrunken log2 fold changes"))
dev.off()

maplot = ggmaplot(resLFC, size = 1,
                  palette = c("#e55c30", "#84206b", "#f6d746"),
                  genenames = as.vector(rownames(resLFC)),
                  legend="top", top = 15,font.label = c("bold", 6),
                  font.legend = c("bold",12), font.main = "bold",
                  xlab = "Log2 Mean Expression",  ylab="Log2 FC")
svg("/home/famgarcia/TFM/data/Testillano col/MV vs PE_HDAC/MAPLOT_ME_HDAC_better.svg")
show(maplot)
dev.off()
```

### Volcano plot {.tabset .tabset-fade .tabset-pills}

For each gene, this plot shows the gene fold change on the x-axis
against the p-value plotted on the y-axis.

Here, we make use of a library called EnhancedVolcano which is available
through Bioconductor and described extensively on its own GitHub page.

First, we are going to "shrink" the log2 fold changes to remove the
noise associated with fold changes coming from genes with low count
levels. Shrinkage of effect size (LFC estimates) is useful for
visualization and ranking of genes. This helps to get more meaningful
log2 fold changes for all genes independently of their expression level.

```{r, warning = FALSE, message = FALSE, echo = FALSE}
# The main function is named after the package
png("/home/famgarcia/TFM/data/Testillano col/MV vs PE_HDAC/VolcanoEnhanced_ME_HDAC.png", res=300, height=30,width=30, units="cm")
EnhancedVolcano(toptable = resLFC,      # We use the shrunken log2 fold change as noise associated with low count genes is removed 
                x = "log2FoldChange",   # Name of the column in resLFC that contains the log2 fold changes
                y = "padj",             # Name of the column in resLFC that contains the p-value
                labSize = 2,
                pCutoff = 0.05,
                FCcutoff = 1,
                xlim = c(-2, 4),
                ylim = c(0,300),
                title="Condition T vs. C", subtitle = NULL,
                lab = rownames(resLFC)) 
dev.off()

# Gather Log-fold change and FDR-corrected pvalues from DESeq2 results

## - Change pvalues to -log10 (1.3 = 0.05)
data <- data.frame(gene = row.names(res),
                   pval = -log10(res$padj), 
                   lfc = res$log2FoldChange)

# Remove any rows that have NA as an entry
data <- na.omit(data)

# Color the points which are up or down
## If fold-change > 0 and pvalue > 1 (Increased significant)
## If fold-change < 0 and pvalue > 1 (Decreased significant)
data <- mutate(data, color = case_when(data$lfc > 1 & data$pval > 1.3 ~ "Increased",
                                       data$lfc < -1 & data$pval > 1.3 ~ "Decreased",
                                       abs(data$lfc) > 1 | data$pval < 1.3 ~ "nonsignificant"))

png("/home/famgarcia/TFM/data/Testillano col/MV vs PE_HDAC/VolcanoFiltering_ME_HDAC.png", res=300, height=20,width=20, units="cm")
# Make a basic ggplot2 object with x-y values
vol <- ggplot(data, aes(x = lfc, y = pval, color = color))

# Add ggplot2 layers
vol +   
  ggtitle(label = "Volcano Plot", subtitle = "Colored by fold-change direction") +
  geom_point(size = 2.5, alpha = 0.8, na.rm = T) +
  scale_color_manual(name = "Directionality",
                     values = c(Increased = "#008B00", Decreased = "#CD4F39", nonsignificant = "darkgray")) +
  theme_bw(base_size = 14) + # change overall theme
  theme(legend.position = "right") + # change the legend
  xlab(expression(log[2]("Log2FoldChange"))) + # Change X-Axis label
  ylab(expression(-log[10]("adjusted p-value"))) + # Change Y-Axis label
  #geom_hline(xintercept = 1.3, colour = "darkgrey") + # Add p-adj value cutoff line
  #geom_hline(yintercept = 1, colour = "darkgrey") + # Add p-adj value cutoff line
  scale_y_continuous(trans = "log1p", oob = scales::squish_infinite, limits = c(0,400))# Scale yaxis due to large p-values

dev.off()
```

### Pheatmap {.tabset .tabset-fade .tabset-pills}

To preserve the order and mappings of the external_gene_names (which are
more informative visually than ensemble_gene_ids) we will need to
perform normalization on dds_new

```{r}
signi <- subset(Res_A_X_total_plot,padj <=0.05)
signi <- subset(signi,log2FoldChange > 1 | log2FoldChange < -1)

normalized_counts <- counts(dds_new, normalized=TRUE)
normalized_counts_mad <- apply(normalized_counts, 1, mad)
normalized_counts <- normalized_counts[order(normalized_counts_mad, decreasing=T), ]

normcounts_A_X <- data.frame(normalized_counts)
normcounts_A_X <- rownames_to_column(normcounts_A_X, var = "external_gene_name")
allsig <- merge(normcounts_A_X, signi,by='external_gene_name')

```

```{r}
sigCounts <- allsig[, 2:7] 

row.names(sigCounts) <- allsig$Row.names

# get the top 20 genes by shrunken log2 fold change
top20 <- order(-abs(resLFC$log2FoldChange))[1:20]
resLFC[top20,]


# negative log-scaled adjusted p-values
log_padj <- -log(resLFC$padj,10)
log_padj[log_padj > 100] <- 100


# get top 20 log fold change genes
top20 <- order(-abs(resLFC$log2FoldChange))[1:100]
df <- data.frame(colData(dds_new)[,"condition"])
	rownames(df) <- colnames(dds_new)
	colnames(df) <- "condition"
	
top20symbols <- subset(Res_A_X_total_plot, ensembl_gene_id %in% rownames(assay(rld)[top20,]))$external_gene_name

df_to_plot <- assay(rld)[top20,]
rownames(df_to_plot) <- top20symbols

png("/home/famgarcia/TFM/data/Testillano col/MV vs PE_HDAC/top100_ME_HDAC.png", res=300, height=40,width=13, units="cm")

pheatmap(
	df_to_plot, 
	cluster_rows=TRUE, 
	show_rownames=TRUE,
	cluster_cols=TRUE,
	annotation_col=df,
	fontsize_row = 8, 
	main= ("Clustering using top 100 log fold change genes")
	)
dev.off()
```
```{r}
resLFC <- lfcShrink(dds, coef="condition_T_vs_VM", type="apeglm")
genelist <- rownames(resLFC)
lfc_df = as.data.frame(resLFC[,2])
lfc_df$ensembl_gene_id = genelist
rownames(lfc_df) <- NULL  
colnames(lfc_df) <-c("lfcShrink","ensembl_gene_id")

final_df <- merge(Res_A_X_total_plot,lfc_df, by = "ensembl_gene_id")

write.table(final_df,"/home/famgarcia/TFM/data/Testillano col/MV vs PE_HDAC/Final_Condition_T_vs_VM_total.tsv", quote=FALSE, sep="\t", row.names = FALSE)

final_df_filtered <- subset(subset(final_df, padj<0.05),abs(log2FoldChange)>1)

write.table(final_df_filtered,"/home/famgarcia/TFM/data/Testillano col/MV vs PE_HDAC/Final_Condition_T_vs_VM_LFC1_padj05.tsv", quote=FALSE, sep="\t", row.names = FALSE)
```
# GO enrichment analysis

Gene Ontology ORA analysis using clusterProfiler

From a list of genes to biological insights

Over Representation Analysis (ORA)

Over Representation Analysis is searching for biological functions or
pathways that are enriched in a list obtained through experimental
studies compared to the complete list of functions/pathways.

The GO resource is divided into 3 main subdomains:

Biological Process (BP): a series of molecular events with a defined
beginning and end relevant for the function of an organism, a cell, etc.
Cellular Component (CC): the part of a cell. Molecular Function (MF):
the enzymatic activites of a gene product.

```{r}
library("biomartr")
library("clusterProfiler")
library("tidyverse")
library("enrichplot")
suppressPackageStartupMessages(library("org.At.tair.db"))
#library("biomaRt")  # only use to remove cache bug

biomartr::organismBM(organism = "Brassica napus")

arabido_attributes = 
  biomartr::organismAttributes("Brassica napus") %>% 
  filter(dataset == "bnapus_eg_gene")
arabido_attributes

attributes_to_retrieve = c("tair_symbol", "entrezgene_id")
```

```{r}
head(Res_A_X_total)
```
The shrinked Log2FoldValues will serve as a reference for subsetting and filtering DEG by 2-fold change of expression (|LFCShrink|>1). This will take into account variation in genes with low LFC values with the prior distribution calculated upon high LFC genes, and thereby have a more conservative view of what genes are significantly differentially expressed.
```{r}
# building the universe!

# we want the correspondence of TAIR/Ensembl symbols with NCBI Entrez gene ids
attributes_to_retrieve = c("uniprotswissprot","entrezgene_id")

resLFC <- lfcShrink(dds, coef="condition_T_vs_VM", type="apeglm")
write.csv(resLFC[which(resLFC$padj<0.05),], "/home/famgarcia/TFM/data/Testillano col/MV vs PE_HDAC/Condition_T_vs_VM_LFCshirnkresults_padj_0.05.csv")
signi <- subset(resLFC,padj <=0.05)
signi <- subset(signi, log2FoldChange >=1 | log2FoldChange <=-1)

# Query the Ensembl API
all_arabidopsis_genes_annotated <- biomartr::biomart(genes = Res_A_X_total[,1], mart= "plants_mart",   dataset = "bnapus_eg_gene",  attributes = attributes_to_retrieve , filters =  "ensembl_gene_id")  
# for compatibility with enrichGO universe
# genes in the universe need to be characters and not integers (Entrez gene id)
all_arabidopsis_genes_annotated$entrezgene_id = as.character(
  all_arabidopsis_genes_annotated$entrezgene_id)
```


```{r}
# retrieving NCBI Entrez gene id for our genes called differential
signi_genes_annotated<- biomartr::biomart(genes = rownames(signi),                  # genes were retrieved using biomartr::getGenome()
                                mart = "plants_mart",                     # marts were selected with biomartr::getMarts()
                                dataset = "bnapus_eg_gene",               # datasets were selected with biomartr::getDatasets()
                                attributes = attributes_to_retrieve,            # attributes were selected with biomartr::getAttributes()
                                filters =   "ensembl_gene_id" )# query key
head(signi_genes_annotated)  


```

```{r}
library(AnnotationHub)
# Retrieve record for Brassica napus OrgDb (NCBI gene ID annotations based on Brassica napus)
ah <- AnnotationHub()
AnnotationHub::query(ah, c("Brassica", "napus"))
# Bnapus OrgDb will be provided with the object AH107389
Bnapus <- ah[["AH107389"]]
```

```{r}
# performing the ORA for Gene Ontology Biological Process class
ora_analysis_bp <- enrichGO(gene = signi_genes_annotated$entrezgene_id, 
                            universe = all_arabidopsis_genes_annotated$entrezgene_id, 
                            OrgDb = Bnapus,  
                            keyType = "ENTREZID",
                            ont = "BP",              # either "BP", "CC" or "MF",
                            pAdjustMethod = "BH",
                            qvalueCutoff = 0.05,
                            readable = TRUE, 
                            pool = FALSE)



ora_analysis_bp_simplify <- clusterProfiler::simplify(ora_analysis_bp, cutoff=0.6, by="p.adjust", select_fun=min)

ora_analysis_bp_df <- ora_analysis_bp@result  %>% tibble::rownames_to_column("GO") %>% filter(p.adjust<=0.05)
readr::write_csv(ora_analysis_bp_df, "Condition_T_vs_C_Bnapus_bp_analysis_res_.csv")

DT::datatable(ora_analysis_bp_df, class = 'cell-border stripe', editable = TRUE,extensions = 'Buttons', options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print')), filter = "top", rownames = F)

barplot(ora_analysis_bp_simplify, 
        drop = TRUE, 
        showCategory = 10, 
        title = "GO Biological Process",
        font.size = 8)
```

```{r}
# performing the ORA for Gene Ontology Biological Process class
ora_analysis_bp <- enrichGO(gene = signi_genes_annotated$entrezgene_id, 
                            universe = all_arabidopsis_genes_annotated$entrezgene_id, 
                            OrgDb = Bnapus,  
                            keyType = "ENTREZID",
                            ont = "CC",              # either "BP", "CC" or "MF",
                            pAdjustMethod = "BH",
                            qvalueCutoff = 0.05,
                            readable = TRUE, 
                            pool = FALSE)



ora_analysis_bp_simplify <- clusterProfiler::simplify(ora_analysis_bp, cutoff=0.6, by="p.adjust", select_fun=min)

ora_analysis_bp_df <- ora_analysis_bp@result  %>% tibble::rownames_to_column("GO") %>% filter(p.adjust<=0.05)
readr::write_csv(ora_analysis_bp_df, "Condition_T_vs_C_Bnapus_cc_analysis_res_.csv")

DT::datatable(ora_analysis_bp_df, class = 'cell-border stripe', editable = TRUE,extensions = 'Buttons', options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print')), filter = "top", rownames = F)

barplot(ora_analysis_bp_simplify, 
        drop = TRUE, 
        showCategory = 10, 
        title = "GO Cellular Component",
        font.size = 8)
```

```{r}
# performing the ORA for Gene Ontology Biological Process class
ora_analysis_bp <- enrichGO(gene = signi_genes_annotated$entrezgene_id, 
                            universe = all_arabidopsis_genes_annotated$entrezgene_id, 
                            OrgDb = Bnapus,  
                            keyType = "ENTREZID",
                            ont = "MF",              # either "BP", "CC" or "MF",
                            pAdjustMethod = "BH",
                            qvalueCutoff = 0.05,
                            readable = TRUE, 
                            pool = FALSE)



ora_analysis_bp_simplify <- clusterProfiler::simplify(ora_analysis_bp, cutoff=0.6, by="p.adjust", select_fun=min)

ora_analysis_bp_df <- ora_analysis_bp@result  %>% tibble::rownames_to_column("GO") %>% filter(p.adjust<=0.05)
readr::write_csv(ora_analysis_bp_df, "Condition_T_vs_C_Bnapus_mf_analysis_res_.csv")

DT::datatable(ora_analysis_bp_df, class = 'cell-border stripe', editable = TRUE,extensions = 'Buttons', options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print')), filter = "top", rownames = F)

barplot(ora_analysis_bp_simplify, 
        drop = TRUE, 
        showCategory = 10, 
        title = "GO Molecular Function",
        font.size = 8)
```