---
title: "get_At_homologs_Ensembl"
author: "Natalia García"
date: '2023-05-12'
output: html_document
---

# Objective : 

getting preprocessed .gmt files and data.frames with genesets that are available in 

* Last PANTHERdb release (v17) - via geneset package (https://github.com/GangLiLab/geneset)
* Ensembl release 56 _Brassica napus_ GO gene categories
* Ensembl release 56 _Arabidopsis thaliana_ GO gene categories from reciprocal ortholog _B.napus_ and _A. thaliana_ gene pairs (Infered through phylogeny and alignment identity values in Ensembl Compara API) in a set of GO categories of interest pertaining to developmental processes, auxin response, transport, synthesis, epigenetic modifications (nucleosome and DNA methylation on cytosine) (see lookup_terms)


Install and load packages

```{r}
BiocManager::install("biomaRt")
library("biomaRt")
library("readxl")
library("geneset")
```

Load GO term list without Bn gene annotations, access TAIR10 Ensembl Plant Arabidopsis Thaliana genes object


```{r}

# create an object for the plants_mart-Ensembl 
ensembl_plants <- useEnsemblGenomes(biomart = "plants_mart")


# create the mart object for both species
arabidopsis_mart <- useMart(biomart = "plants_mart", host = "https://plants.ensembl.org", dataset = "athaliana_eg_gene")
bnapus_mart <- useMart(biomart = "plants_mart", host = "https://plants.ensembl.org", dataset = "bnapus_eg_gene"	)


# get homology information
hom_At_Bn_list <- getLDS(attributes = c("ensembl_gene_id","external_gene_name","athaliana_eg_homolog_orthology_type", "athaliana_eg_homolog_orthology_confidence", "athaliana_eg_homolog_perc_id_r1"),
              mart = bnapus_mart,
              attributesL = c("ensembl_gene_id","external_gene_name","description","bnapus_eg_homolog_orthology_type", "bnapus_eg_homolog_orthology_confidence", "bnapus_eg_homolog_perc_id_r1"),
              martL = arabidopsis_mart
              )

```
```{r}
getBM(attributes = c("athaliana_eg_homolog_ensembl_gene", "athaliana_eg_homolog_associated_gene_name", "ensembl_gene_id","external_gene_name"))
```

Get TAIR GO terms subgroup from our selected subset of GO terms referring to auxin transport, response, metabolism, epigenetic processes, cell ploriferation and developmental process.

```{r setup, include=FALSE}

# Get GO terms present in arabidopsis but not present in Brassica for homolog search and completion of gene sets.

go_terms_at <- getBM(mart=arabidopsis_mart, attributes=c("go_id", "name_1006", "namespace_1003"))

lookup_terms = c("GO:0051093","GO:0051094","GO:0090644","GO:0030583","GO:0003006","GO:0098727","GO:0022611","GO:0048532","GO:0060033","GO:0021700","GO:0031128","GO:0009838","GO:0043696","GO:0043934","GO:0048589","GO:0048646","GO:0048869","GO:0090693","GO:0097737","GO:0016049","GO:0090354","GO:0080024","GO:0090356","GO:0090355","GO:0010249","GO:0009683","GO:0080161","GO:0071365","GO:0080026","GO:0009721","GO:0016577","GO:0061647","GO:0000412","GO:1990453","GO:0006337","GO:0090124","GO:0010424","GO:0010425","GO:0010426")


go_gset <- getBM(mart=arabidopsis_mart, attributes=c("go_id", "name_1006", "ensembl_gene_id", "namespace_1003"))

go_id_lookup_geneset = go_gset[which(go_gset[,1] %in% lookup_terms),]

go_gset_at_bp <-go_id_lookup_geneset[which(go_id_lookup_geneset[,4] == "biological_process"),c(1,2,3)]

```

Get the corresponding high confidence "reciprocal ortholog" gene pair list

```{r}
# Merge A. thaliana genesets of interest with ortholog list

ortholog_gs_lookup = merge(hom_At_Bn_list, go_id_lookup_geneset, by.x="Gene.stable.ID.1", by.y="ensembl_gene_id")

# get high confidence orthologs both for Arabidopsis thaliana and Brassica napus

hc_ortholog_lookup = ortholog_gs_lookup[which(ortholog_gs_lookup[,5] == 1 & ortholog_gs_lookup[,10] == 1) ,]

# Getting gs - orthologues by ENA identifiers

hc_ortholog_lookup_name <- hc_ortholog_lookup[,c(12,3)]
colnames(hc_ortholog_lookup_name) <- c("bp", "gene")
hc_ortholog_lookup_term <- hc_ortholog_lookup[,c(12,13)]
colnames(hc_ortholog_lookup_term) <- c("bp", "name")

# Labelling gs
#hc_ortholog_lookup_eid$go_id = lapply(hc_ortholog_lookup_eid$go_id, function(x){paste(x,"_o",sep="")})
#hc_ortholog_lookup_name$go_id = lapply(hc_ortholog_lookup_name$go_id, function(x){paste(x,"_o",sep="")})

```


Last step : get _Brassica napus_ Ensembl Release 56 and PANTHERdb available genesets

```{r}
library("data.table")



############## ENSEMBL - UniprotKB curated GO BASED annotations
# get and save brassica geneset

go_gset_bn <- getBM(mart=bnapus_mart, attributes=c("go_id","namespace_1003","ensembl_gene_id", "name_1006"))
go_gset_bn_pr <- go_gset_bn[which(go_gset_bn[,1] != "" ), ]

# Separate by

# Biological process
go_gset_bn_bp <-go_gset_bn_pr[which(go_gset_bn_pr[,2] == "biological_process"),]


# Molecular function
go_gset_bn_mf <-go_gset_bn_pr[which(go_gset_bn_pr[,2] == "molecular_function"),]


# Cellular Component
go_gset_bn_cc <-go_gset_bn_pr[which(go_gset_bn_pr[,2] == "cellular_component"),]



# Map and merge to ENA identifiers

map_name_ensembl <- getBM(mart=bnapus_mart, attributes=c("ensembl_gene_id", "external_gene_name"))
map_name_ensembl <- map_name_ensembl[which(map_name_ensembl[,2] != "" ),]

gset_bn_bp_name <- merge(go_gset_bn_bp,map_name_ensembl, by="ensembl_gene_id")
gset_bn_bp_gs <- gset_bn_bp_name[,c(2,5)]
colnames(gset_bn_bp_gs) <- c("bp", "gene")

gset_bn_bp_term <- gset_bn_bp_name[,c(2,4)]
colnames(gset_bn_bp_term) <- c("bp", "name")


gset_bn_mf_name <- merge(go_gset_bn_mf,map_name_ensembl, by="ensembl_gene_id")
gset_bn_mf_gs <- gset_bn_mf_name[,c(2,5)]
colnames(gset_bn_mf_gs) <- c("mf", "gene")

gset_bn_mf_term <- gset_bn_mf_name[,c(2,4)]
colnames(gset_bn_mf_term) <- c("mf", "name")


gset_bn_cc_name <- merge(go_gset_bn_cc,map_name_ensembl, by="ensembl_gene_id")
gset_bn_cc_gs <- gset_bn_cc_name[,c(2,5)]
colnames(gset_bn_cc_gs) <- c("cc", "gene")

gset_bn_cc_term <- gset_bn_cc_name[,c(2,4)]
colnames(gset_bn_cc_term) <- c("cc", "name")



############## ENSEMBL - UniprotKB curated GO BASED annotations

pathways_Panther_BP <- geneset::getGO(org = "Brassica napus", ont = "bp")
pathways_Panther_MF <- geneset::getGO(org = "Brassica napus", ont = "mf")
pathways_Panther_CC <- geneset::getGO(org = "Brassica napus", ont = "cc")

panther_gset_BP <- pathways_Panther_BP$geneset
panther_terms_BP <-pathways_Panther_BP$geneset_name
colnames(panther_terms_BP) <- c("bp","name")


panther_gset_MF <- pathways_Panther_MF$geneset
panther_terms_MF <-pathways_Panther_MF$geneset_name
colnames(panther_terms_MF) <- c("mf","name")


panther_gset_CC <- pathways_Panther_CC$geneset
panther_terms_CC <-pathways_Panther_CC$geneset_name
colnames(panther_terms_CC) <- c("cc","name")



############## Bind all annnotations to Arabidopsis selected gs of interest reciprocal ortholog list

full_genesets_BP <- unique(rbind(hc_ortholog_lookup_name, rbind(gset_bn_bp_gs,panther_gset_BP)))
full_genesets_MF <- unique(rbind(gset_bn_mf_gs,panther_gset_MF))
full_genesets_CC <- unique(rbind(gset_bn_cc_gs,panther_gset_CC))


full_terms_BP <- unique(rbind(hc_ortholog_lookup_term, rbind(gset_bn_bp_term,panther_terms_BP)))
full_terms_MF <- unique(rbind(gset_bn_mf_term,panther_terms_MF))
full_terms_CC <- unique(rbind(gset_bn_cc_term,panther_terms_CC))

```

Save in csv format
```{r}
library("data.table")


fwrite(full_genesets_BP, 'C:\\Users\\naata\\Documents\\MASTER\\BioSynth\\synthetic_env\\GSEA\\full_bp_genesets.csv', row.names=FALSE)
fwrite(full_genesets_MF, 'C:\\Users\\naata\\Documents\\MASTER\\BioSynth\\synthetic_env\\GSEA\\full_mf_genesets.csv', row.names=FALSE)
fwrite(full_genesets_CC, 'C:\\Users\\naata\\Documents\\MASTER\\BioSynth\\synthetic_env\\GSEA\\full_cc_genesets.csv', row.names=FALSE)

fwrite(full_terms_BP, 'C:\\Users\\naata\\Documents\\MASTER\\BioSynth\\synthetic_env\\GSEA\\full_bp_terms.csv', row.names=FALSE)
fwrite(full_terms_MF, 'C:\\Users\\naata\\Documents\\MASTER\\BioSynth\\synthetic_env\\GSEA\\full_mf_terms.csv', row.names=FALSE)
fwrite(full_terms_CC, 'C:\\Users\\naata\\Documents\\MASTER\\BioSynth\\synthetic_env\\GSEA\\full_cc_terms.csv', row.names=FALSE)

```
Preprocess and save into GMT files

```{r}
library(stringr)
library(readr)
library(dplyr)
library(tidyr)

# Save as multilevel list tidy format through pivot_longer and pivot wider

# Biological Process
ml_BP_GO <- full_genesets_BP %>%
  pivot_longer(gene, names_to = "Var", names_prefix = "Var") %>% 
  pivot_wider(names_from = bp, values_from = value, names_sep = "", values_fn = list)

# Cellular Component
ml_CC_GO <- full_genesets_MF %>%
  pivot_longer(gene, names_to = "Var", names_prefix = "Var") %>% 
  pivot_wider(names_from = mf, values_from = value, names_sep = "", values_fn = list)

# Molecular Function
ml_MF_GO <- full_genesets_CC %>%
  pivot_longer(gene, names_to = "Var", names_prefix = "Var") %>% 
  pivot_wider(names_from =cc, values_from = value, names_sep = "", values_fn = list)


```


```{r}
# Handling pivoted multilevel lists and converting into gmt file format

# GO_BP
listindex <- seq(ml_BP_GO)
pathwayindex <- names(ml_BP_GO)

file.create("C:\\Users\\naata\\Documents\\MASTER\\BioSynth\\synthetic_env\\Bna_GO_BP_formatted.gmt")
invisible(lapply(listindex[2:length(listindex)], function(i){
  cat(names(ml_BP_GO)[i], "N/A", lapply(ml_BP_GO[[i]], paste, collapse = "\t")[[1]], "\n", sep="\t", file="C:\\Users\\naata\\Documents\\MASTER\\BioSynth\\synthetic_env\\Bna_GO_BP_formatted.gmt", append=TRUE)}))

# GO_CC
listindex <- seq(ml_CC_GO)
pathwayindex <- names(ml_CC_GO)

file.create("C:\\Users\\naata\\Documents\\MASTER\\BioSynth\\synthetic_env\\Bna_GO_CC_formatted.gmt")
invisible(lapply(listindex[2:length(listindex)], function(i){
  cat(names(ml_CC_GO)[i], "N/A", lapply(ml_CC_GO[[i]], paste, collapse = "\t")[[1]], "\n", sep="\t", file="C:\\Users\\naata\\Documents\\MASTER\\BioSynth\\synthetic_env\\Bna_GO_CC_formatted.gmt", append=TRUE)}))

# GO_CC
listindex <- seq(ml_MF_GO)
pathwayindex <- names(ml_MF_GO)

file.create("C:\\Users\\naata\\Documents\\MASTER\\BioSynth\\synthetic_env\\Bna_GO_MF_formatted.gmt")
invisible(lapply(listindex[2:length(listindex)], function(i){
  cat(names(ml_MF_GO)[i], "N/A", lapply(ml_MF_GO[[i]], paste, collapse = "\t")[[1]], "\n", sep="\t", file="C:\\Users\\naata\\Documents\\MASTER\\BioSynth\\synthetic_env\\Bna_GO_MF_formatted.gmt", append=TRUE)}))


```



# GET KEGG LIST from 


## Preprocess KEGG database gene names - Ensembl ID to external gene name mappings

```{r}

# Load libraries
suppressPackageStartupMessages(library("tidyverse"))
suppressPackageStartupMessages(library("biomaRt"))
suppressPackageStartupMessages(library("biomartr"))

pathways_KEGG <- geneset::getKEGG(org = "bna",category = "pathway")

KEGG_table_preprocess = read.table("C:\\Users\\naata\\Documents\\MASTER\\BioSynth\\synthetic_env\\Bna_ALL.KEGG.txt", header=TRUE, sep="\t")

KEGG <- KEGG_table_preprocess[,1:2]


# get gene names and transcript lengths when they exist
ann <- getBM(filter="ensembl_gene_id",value=KEGG$Gene, attributes=c("ensembl_gene_id","external_gene_name"), mart=bnapus_mart, useCache = FALSE)

KEGG <-merge(KEGG,ann,by.x="Gene", by.y="ensembl_gene_id",all.x = FALSE, all.y = TRUE)

# Subset only external genes
KEGG_term2gene <- KEGG[,c(2,3)]
colnames(KEGG_term2gene) <- c("term","gene")

KEGG_term2name <- KEGG[,c(2,2)]
colnames(KEGG_term2name) <- c("term","name")

write.csv(KEGG_term2gene, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\tfm\\KEGG_gs_pairs.csv", row.names =FALSE)
write.csv(KEGG_term2name, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\tfm\\KEGG_gs_descriptions.csv", row.names =FALSE)

# KEGG Pathways
ml_KEGG <- KEGG %>%
  pivot_longer(external_gene_name, names_to = "Var", names_prefix = "Var") %>% 
  pivot_wider(names_from = Gene.set.name, values_from = value, names_sep = "", values_fn = list)


# Save KEGG
listindex <- seq(ml_KEGG)
pathwayindex <- names(ml_KEGG)

file.create("C:\\Users\\naata\\Documents\\MASTER\\BioSynth\\synthetic_env\\Bna_KEGG_formatted.gmt")
invisible(lapply(listindex[2:length(listindex)], function(i){
  cat(names(ml_KEGG)[i], "N/A", lapply(ml_KEGG[[i]], paste, collapse = "\t")[[1]], "\n", sep="\t", file="C:\\Users\\naata\\Documents\\MASTER\\BioSynth\\synthetic_env\\Bna_KEGG_formatted.gmt", append=TRUE)}))


```





# In case Entrez ID gene categories need to be accessed - NCBI AnnotationHub object (only Brassica napus annotation available)

```{r}

#hc_ortholog_lookup_eid <- hc_ortholog_lookup[,c(12,2)]
colnames(hc_ortholog_lookup_eid) <- c("bp","gene")

# Map and merge to Entrez ID identifiers (less mappings observed)

map_eid_ensembl <- getBM(mart=bnapus_mart, attributes=c("ensembl_gene_id", "entrezgene_id"))

map_eid_ensembl <- map_eid_ensembl[which(map_eid_ensembl[,2] != "" ),]

library("AnnotationForge")
library("AnnotationDbi")
library("AnnotationHub")
library("OrgDB")

# Making annotation package
ah <- AnnotationHub()

#Ensembl 56
query(ah, c( "OrgDb", "Brassica napus"))

org.Brassicanapus.eg.db <- ah[["AH107389"]]

egid <- keys(org.Brassicanapus.eg.db, "ENTREZID")

db_gs_ncbi <- select(org.Brassicanapus.eg.db, egid, c("GENENAME", "GO"), "ENTREZID")

gs_ncbi_filtered <- db_gs_ncbi[which(db_gs_ncbi[,3] != "<NA>"),c(1,3)]

```


```{r}
DEG_Stress <-read_csv("C:\\Users\\naata\\MASTER\\Enrichment\\Final_condition_PErep_vs_VMrep_padj0.05_woPErep3_LFC1filtered.csv")
DEG_Stress_gn<- merge(DEG_Stress, map_name_ensembl, by="ensembl_gene_id")

fwrite(DEG_Stress_gn, "C:\\Users\\naata\\MASTER\\Enrichment\\Final_condition_PErep_vs_VMrep_padj0.05_woPErep3_LFC1filtered.csv", row.names = FALSE)
```

