---
title: "GSEA_v2"
author: "Natalia García"
date: '2023-05-16'
output: html_document
---
In this script we will

* Preprocess Gene Ontology (GO) gene set - "external gene name" (European Nucleoide Archive identifiers) annotations retrieved from BiomaRt API programmatic access - Ensembl Plants release 56 - _Brassica napus_ genome + labelled GO _Arabidopsis thaliana_  annotations from Reciprocal orthologous gene pairs Compara API in a set of selected categories of interest (See Annex I of Master thesis for more details ). Annotations are intended to be more comprehensive and specific than `PANTHERdb` or `PlantGSAD`, and Ensembl transcript centric contrary to those available in NCBI AnnotationHub R package.

* Conduct a FCS Gene Enrichment Analysis (GSEA) using the preranked GSEA algorithm with the package `genekitr`. 

# Loading libraries

```{r setup, include=FALSE}
suppressPackageStartupMessages(library("pathview"))
suppressPackageStartupMessages(library("readxl"))
suppressPackageStartupMessages(library("clusterProfiler"))
suppressPackageStartupMessages(library("Rgraphviz"))
suppressPackageStartupMessages(library("graph"))
library(tidyverse)
library(geneset)
library(genekitr)
library(data.table)
suppressPackageStartupMessages(library("KEGGgraph"))
suppressWarnings(suppressPackageStartupMessages(library("fgsea")))
```

# Loading data

A collection of gene - DeSeq2 Wald statistic pair values is used in a `.rnk` file, sorted by the latter statistic representing gene expression levels

(Enter wald statistic formula : (/Zscore))

```{r cars}
ranks_HDAC <- read.csv("C:\\Users\\naata\\Documents\\MASTER\\BioSynth\\synthetic_env\\GSEA\\bna_HDAC_deseq2stat.rnk",
                    header=FALSE, colClasses = c("character", "numeric"))


ranks_Stress <- read.csv("C:\\Users\\naata\\Documents\\MASTER\\BioSynth\\synthetic_env\\GSEA\\bna_MVvsPE_deseq2stat.rnk",
                    header=FALSE, colClasses = c("character", "numeric"))

# We'll extract the background genes - the genes detected with over log10(counts)>1

background_ALL_gl_HDAC <- ranks_HDAC[,1]
background_ALL_gl_Stress <- ranks_Stress[,1]

background_UP_gl_HDAC <- ranks_HDAC[which(ranks_HDAC[,2] > 0), 1]
background_UP_gl_Stress <- ranks_Stress[which(ranks_Stress[,2] > 0), 1]

background_DOWN_gl_HDAC <- ranks_HDAC[which(ranks_HDAC[,2] < 0), 1]
background_DOWN_gl_Stress <- ranks_Stress[which(ranks_Stress[,2] < 0), 1]

```

## Preprocessing data

```{r pressure, echo=FALSE}


# Remove duplicates (transcript isoform with more differential expression will be conserved from the same gene)
ranks_HDAC <- ranks_HDAC[!duplicated(ranks_HDAC[,1]),]
ranks_Stress <- ranks_Stress[!duplicated(ranks_Stress[,1]),]

# Labelling columns (gene_name "ID" and statistic "t")
colnames(ranks_HDAC) <-c("ID","t")
colnames(ranks_Stress) <-c("ID","t")

# Sorting and creating ordered gene ranks column
ranks_HDAC <- ranks_HDAC %>% mutate(rank = rank(t,  ties.method = "random"))
ranks_HDAC <- ranks_HDAC[order(ranks_HDAC$rank, decreasing=TRUE), ]

ranks_Stress <- ranks_Stress %>% mutate(rank = rank(t,  ties.method = "random"))
ranks_Stress <- ranks_Stress[order(ranks_Stress$rank, decreasing=TRUE), ]

# Converting to set of lists
ranks_HDAC <- setNames(ranks_HDAC$t, ranks_HDAC$ID)
ranks_Stress <- setNames(ranks_Stress$t, ranks_Stress$ID)


```


# Loading pathways and background gene set

```{r}
# Background gene set with detected gene expression - Used in Overrepresentation analysis
disreg_genes<-read_table("C:\\Users\\naata\\MASTER\\Enrichment\\Filtered_DEG_LFC1_padj05_results_total.xlsx")

# Biomart access - get mappings, and disreg gene list (UP, DOWN, ALL)

```


# Loading gene set categories obtained in "get_gs_homolog_gs.Rmd"
```{r}

# Biological Process

pathways_BP_gs <-read.csv('C:\\Users\\naata\\Documents\\MASTER\\BioSynth\\synthetic_env\\GSEA\\full_bp_genesets.csv')
pathways_MF_gs <-read.csv('C:\\Users\\naata\\Documents\\MASTER\\BioSynth\\synthetic_env\\GSEA\\full_mf_genesets.csv')
pathways_CC_gs <-read.csv('C:\\Users\\naata\\Documents\\MASTER\\BioSynth\\synthetic_env\\GSEA\\full_cc_genesets.csv')

pathways_BP_terms <-read.csv('C:\\Users\\naata\\Documents\\MASTER\\BioSynth\\synthetic_env\\GSEA\\full_bp_terms.csv')
pathways_MF_terms <-read.csv('C:\\Users\\naata\\Documents\\MASTER\\BioSynth\\synthetic_env\\GSEA\\full_mf_terms.csv')
pathways_CC_terms <-read.csv('C:\\Users\\naata\\Documents\\MASTER\\BioSynth\\synthetic_env\\GSEA\\full_cc_terms.csv')

```

# Biological process GSEA

```{r}
library("clusterProfiler")
library(stringr)

attr<- c("ID","Description","setSize","enrichmentScore","NES","pvalue","p.adjust","qvalue","rank","leading_edge","geneID", "geneID_symbol", "Count","GeneRatio")

gsea_HDAC_o <- clusterProfiler::GSEA(geneList = ranks_HDAC, minGSSize=2, eps=0.0, pvalueCutoff=1, pAdjustMethod="BH", gson=NULL, TERM2GENE = pathways_BP_gs, TERM2NAME = pathways_BP_terms, seed=TRUE, by="fgsea")
gsea_HDAC<-gsea_HDAC_o[,1:11]

gsea_Stress_o <- clusterProfiler::GSEA(geneList = ranks_Stress, minGSSize=2, eps=0.0, pvalueCutoff=1, pAdjustMethod="BH", gson=NULL, TERM2GENE = pathways_BP_gs, TERM2NAME = pathways_BP_terms, seed=TRUE, by="fgsea")
gsea_Stress<-gsea_Stress_o[,1:11]

count_HDAC <- c()
gsea_HDAC$geneID_symbol = gsea_HDAC$core_enrichment
for (x in gsea_HDAC$core_enrichment){count_HDAC<-c(count_HDAC, as.numeric(str_count(x, "/")+1))}
gsea_HDAC$Count <- count_HDAC
gsea_HDAC$GeneRatio <- gsea_HDAC$Count/gsea_HDAC$setSize
names(gsea_HDAC) <- attr

count_Stress <- c()
gsea_Stress$geneID_symbol = gsea_Stress$core_enrichment
for (x in gsea_Stress$core_enrichment){count_Stress<-c(count_Stress, as.numeric(str_count(x, "/")+1))}
gsea_Stress$Count <- count_Stress
gsea_Stress$GeneRatio <- gsea_Stress$Count/gsea_Stress$setSize
names(gsea_Stress) <- attr


```

Preprocessing genekitr for plotting

```{r}

# LFC

lfc_val_HDAC <- read.csv("C:\\Users\\naata\\Documents\\MASTER\\BioSynth\\synthetic_env\\GSEA\\bna_HDAC_deseq2stat.rnk",
                    header=FALSE, colClasses = c("character", "numeric"))
lfc_val_Stress <- read.csv("C:\\Users\\naata\\Documents\\MASTER\\BioSynth\\synthetic_env\\GSEA\\bna_MVvsPE_deseq2stat.rnk",
                    header=FALSE, colClasses = c("character", "numeric"))

colnames(lfc_val_HDAC) <- c("ID", "logfc")
colnames(lfc_val_Stress) <- c("ID", "logfc")

# METADATA

org_df <- data.frame(org="bnapus")
exp_df <- data.frame(exponent=1)

# GSEA result

gsea_HDAC_gkit <- as.enrichdat(gsea_HDAC[1:11])
gsea_Stress_gkit <- as.enrichdat(gsea_Stress[1:11])

BP_HDAC_list = list()
BP_HDAC_list[['gsea_df']] <- gsea_HDAC
BP_HDAC_list[['genelist']] <- lfc_val_HDAC
BP_HDAC_list[['geneset']] <- pathways_BP_gs[!(pathways_BP_gs[,1] %in% c(child_s_go[,1],parent_s_go[,1])),]
BP_HDAC_list[['exponent']] <- exp_df
BP_HDAC_list[['org']] <- org_df


BP_Stress_list = list()
BP_Stress_list[['gsea_df']] <- gsea_Stress
BP_Stress_list[['genelist']] <- lfc_val_Stress
BP_Stress_list[['geneset']] <- pathways_BP_gs[!(pathways_BP_gs[,1] %in% c(child_s_go[,1],parent_s_go[,1])),]
BP_Stress_list[['exponent']] <- exp_df
BP_Stress_list[['org']] <- org_df

```

# LOAD genesets of interest from Excel file
```{r}
child_s_go <- read_excel("C:\\Users\\naata\\Documents\\MASTER\\TFM\\tfm\\GO_to_lookup_homologuesAt.xlsx", sheet="CHILD_TERMS_ALL")

parent_s_go <- read_excel("C:\\Users\\naata\\Documents\\MASTER\\TFM\\tfm\\GO_to_lookup_homologuesAt.xlsx", sheet="MAIN_TERMS")

colnames(child_s_go) <- c("ID", "name", "parent")
colnames(parent_s_go) <- c("ID", "name")


write.xlsx(BP_HDAC_list, "C://Users//naata//Desktop//reprex_plotGSEA.xlsx")
```


```{r}
# merge gsea_Results GO + child/parent terms
pathways = as.data.frame(gsea_HDAC[which(gsea_HDAC[,1] %in% parent_s_go[,1]$ID),1])
colnames(pathways)="ID"
genekitr::plotGSEA(BP_HDAC_list, plot_type = "classic")

# Error replacement calcscore(geneset genelist, pathway, exponent, fortify=TRUE)
# see docs
# or 
# keep preprocessing data?
```
```{r}
genekitr::plotGSEA(BP_HDAC_list, plot_type = "classic", show_pathway = "GO:0040029")
```


# OVERREPRESENTATION ANALYSIS

# Import results from DeSeq2 analysis
```{r}
DEG_HDAC<-read_excel("C:\\Users\\naata\\MASTER\\Enrichment\\Filtered_DEG_LFC1_padj05_results_total.xlsx")
DEG_Stress <-read_csv("C:\\Users\\naata\\MASTER\\Enrichment\\Final_condition_PErep_vs_VMrep_padj0.05_woPErep3_LFC1filtered.csv")

DEG_ALL_HDAC <- DEG_HDAC$external_gene_name
DEG_ALL_Stress <- DEG_Stress$external_gene_name

DEG_UP_HDAC <- DEG_HDAC[which(DEG_HDAC$log2FoldChange > 0),]$external_gene_name
DEG_UP_Stress <- DEG_Stress[which(DEG_Stress$log2FoldChange > 0),]$external_gene_name


DEG_DOWN_HDAC <- DEG_HDAC[which(DEG_HDAC$log2FoldChange < 0),]$external_gene_name
DEG_DOWN_Stress <- DEG_Stress[which(DEG_Stress$log2FoldChange < 0),]$external_gene_name

```

# All genes
```{r}
# Get background genes : genes with detected minimal expression
colnames(background_ALL_gl_HDAC) = "gene"
bckgr_geneset_HDAC <- merge(background_ALL_gl_HDAC, pathways_BP_gs, by="gene")

colnames(background_ALL_gl_Stress) = "gene"
bckgr_geneset_Stress <- merge(background_ALL_gl_Stress, pathways_BP_gs, by="gene")

ORA_HDAC_ALL <- clusterProfiler::enricher(geneList = DEG_ALL_HDAC, minGSSize=2, eps=0.0, pvalueCutoff=1, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_HDAC, TERM2NAME = pathways_BP_terms, seed=TRUE, by="fgsea")
ORA_Stress <- clusterProfiler::enricher(geneList = DEG_ALL_Stress, minGSSize=2, eps=0.0, pvalueCutoff=1, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_Stress, TERM2NAME = pathways_BP_terms, seed=TRUE, by="fgsea")


```

# Up-regulated genes

```{r}
library("clusterProfiler")
colnames(background_ALL_gl_HDAC) = "gene"
bckgr_geneset_HDAC <- merge(background_UP_gl_HDAC, pathways_BP_gs, by="gene")

colnames(background_ALL_gl_Stress) = "gene"
bckgr_geneset_Stress <- merge(background_UP_gl_Stress, pathways_BP_gs, by="gene")

ORA_HDAC_ALL <- clusterProfiler::enricher(geneList = DEG_UP_HDAC, minGSSize=2, eps=0.0, pvalueCutoff=1, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_HDAC, TERM2NAME = pathways_BP_terms, seed=TRUE, by="fgsea")
ORA_Stress <- clusterProfiler::enricher(geneList = DEG_UP_Stress, minGSSize=2, eps=0.0, pvalueCutoff=1, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_Stress, TERM2NAME = pathways_BP_terms, seed=TRUE, by="fgsea")


```

# Down-regulated genes

```{r}
library("clusterProfiler")
colnames(background_ALL_gl_HDAC) = "gene"
bckgr_geneset_HDAC <- merge(background_DOWN_gl_HDAC, pathways_BP_gs, by="gene")

colnames(background_ALL_gl_Stress) = "gene"
bckgr_geneset_Stress <- merge(background_DOWN_gl_Stress, pathways_BP_gs, by="gene")

ORA_HDAC_ALL <- clusterProfiler::enricher(geneList = DEG_DOWN_HDAC, minGSSize=2, eps=0.0, pvalueCutoff=1, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_HDAC, TERM2NAME = pathways_BP_terms, seed=TRUE, by="fgsea")
ORA_Stress <- clusterProfiler::enricher(geneList = DEG_DOWN_Stress, minGSSize=2, eps=0.0, pvalueCutoff=1, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_Stress, TERM2NAME = pathways_BP_terms, seed=TRUE, by="fgsea")


```
