---
title: "ORA_analysis"
author: "Natalia García"
date: '2023-05-17'
output: html_document
---

Overrepresentation analysis of genesets


* Conduct an overrepresentation analysis hyper-geometric - exact Fisher test with FDR correction for Gene Enrichment Analysis using clusterProfiler

Collapse pathways
Save results
Visualize lookup genes

conduct for PLANTGSAD (TF, Cyc, KEGG)

# Loading libraries

```{r setup, include=FALSE}
suppressPackageStartupMessages(library("pathview"))
suppressPackageStartupMessages(library("readxl"))
suppressPackageStartupMessages(library("clusterProfiler"))
suppressPackageStartupMessages(library("Rgraphviz"))
suppressPackageStartupMessages(library("graph"))
library(tidyverse)
library(geneset)
library(genekitr)
library(data.table)
library("rrvgo")
library("igraph")
library("ggraph")
suppressPackageStartupMessages(library("KEGGgraph"))
suppressWarnings(suppressPackageStartupMessages(library("fgsea")))
```

Load available AnnotationHub Brassica napus annotation object to construct an Orgdb object for rrvgo to perform on

```{r}

library("AnnotationForge")
library("AnnotationDbi")
library("AnnotationHub")

# Making annotation package
ah <- AnnotationHub()

#Ensembl 56
query(ah, c( "OrgDb", "Brassica napus"))

org.Brassicanapus.eg.db <- ah[["AH107389"]]


```

# Loading data

A collection of gene - DeSeq2 Wald statistic pair values is used in a `.rnk` file, sorted by the latter statistic representing gene expression levels

(Enter wald statistic formula : (/Zscore))

```{r cars}

ranks_HDAC_Stress <- read.csv("C:\\Users\\naata\\MASTER\\Enrichment\\Rank_T_vs_VM.rnk",
                              header=FALSE, colClasses = c("character", "numeric"))

ranks_HDAC <- read.csv("C:\\Users\\naata\\Documents\\MASTER\\BioSynth\\synthetic_env\\GSEA\\bna_HDAC_deseq2stat.rnk",
                    header=FALSE, colClasses = c("character", "numeric"))


ranks_Stress <- read.csv("C:\\Users\\naata\\Documents\\MASTER\\BioSynth\\synthetic_env\\GSEA\\bna_MVvsPE_deseq2stat.rnk",
                    header=FALSE, colClasses = c("character", "numeric"))

# We'll extract the background genes - the genes detected with over log10(counts)>1
# Remove duplicates (transcript isoform with more differential expression will be conserved from the same gene)
ranks_HDAC <- ranks_HDAC[!duplicated(ranks_HDAC[,1]),]
ranks_Stress <- ranks_Stress[!duplicated(ranks_Stress[,1]),]
ranks_HDAC_Stress <- ranks_HDAC_Stress[!duplicated(ranks_HDAC_Stress[,1]),]


background_ALL_gl_HDAC <- as.data.frame(ranks_HDAC[,1])
background_ALL_gl_Stress <- as.data.frame(ranks_Stress[,1])
background_ALL_gl_HDAC_Stress <- as.data.frame(ranks_HDAC_Stress[,1])

background_UP_gl_HDAC <- as.data.frame(ranks_HDAC[which(ranks_HDAC[,2] > 0), 1])
background_UP_gl_Stress <- as.data.frame(ranks_Stress[which(ranks_Stress[,2] > 0), 1])
background_UP_gl_HDAC_Stress <- as.data.frame(ranks_HDAC_Stress[which(ranks_HDAC_Stress[,2] > 0), 1])

background_DOWN_gl_HDAC <- as.data.frame(ranks_HDAC[which(ranks_HDAC[,2] < 0), 1])
background_DOWN_gl_Stress <- as.data.frame(ranks_Stress[which(ranks_Stress[,2] < 0), 1])
background_DOWN_gl_HDAC_Stress <- as.data.frame(ranks_HDAC_Stress[which(ranks_HDAC_Stress[,2] < 0), 1])

```

```{r}

hormone_terms = read_excel("C:\\Users\\naata\\Documents\\MASTER\\TFM\\tfm\\final_list_GO_lookup.xlsx", sheet="hormonal_screening")
epi_terms = read_excel("C:\\Users\\naata\\Documents\\MASTER\\TFM\\tfm\\final_list_GO_lookup.xlsx", sheet="epigenetic_screening")

```



# Loading pathways and background gene set


# Loading gene set categories obtained in "get_gs_homolog_gs.Rmd"
```{r}

# Biological Process

ml_BP_GO <- readRDS("C:\\Users\\naata\\Documents\\MASTER\\BioSynth\\synthetic_env\\GSEA\\full_bp_genesets_noorth.rds")
ml_CC_GO <- readRDS("C:\\Users\\naata\\Documents\\MASTER\\BioSynth\\synthetic_env\\GSEA\\full_cc_genesets.rds")
ml_MF_GO <- readRDS("C:\\Users\\naata\\Documents\\MASTER\\BioSynth\\synthetic_env\\GSEA\\full_mf_genesets.rds")

# Handling pivoted multilevel lists and converting into gmt file format

# GO_BP
listindex <- seq(ml_BP_GO)
pathwayindex <- names(ml_BP_GO)

# Pass to df format
tmp_list <- lapply(listindex, function(i){
                        gl = list(ml_BP_GO[[i]])[[1]]
                        cbind(rep(names(ml_BP_GO)[i],length(gl)),gl)})

pathways_BP_gs<-do.call(rbind,  tmp_list)
colnames(pathways_BP_gs)=c("bp", "gene")


# GO_CC
listindex <- seq(ml_CC_GO)
pathwayindex <- names(ml_CC_GO)

# Pass to df format
tmp_list <- lapply(listindex, function(i){
                        gl = list(ml_CC_GO[[i]])[[1]]
                        cbind(rep(names(ml_CC_GO)[i],length(gl)),gl)})

# Write to gmt
pathways_CC_gs<-do.call(rbind,  tmp_list)
colnames(pathways_CC_gs)=c("cc", "gene")


# GO_MF
listindex <- seq(ml_MF_GO)
pathwayindex <- names(ml_MF_GO)

# Pass to df format
tmp_list <- lapply(listindex, function(i){
                        gl = list(ml_MF_GO[[i]])[[1]]
                        cbind(rep(names(ml_MF_GO)[i],length(gl)),gl)})

pathways_MF_gs<-do.call(rbind,  tmp_list)
colnames(pathways_MF_gs)=c("mf", "gene")


# KEGG
pathways_KEGG_gs <- read.csv('C:\\Users\\naata\\Documents\\MASTER\\TFM\\tfm\\KEGG_gs_pairs.csv')


pathways_BP_terms <-read.csv('C:\\Users\\naata\\Documents\\MASTER\\BioSynth\\synthetic_env\\GSEA\\full_bp_terms.csv')
pathways_MF_terms <-read.csv('C:\\Users\\naata\\Documents\\MASTER\\BioSynth\\synthetic_env\\GSEA\\full_mf_terms.csv')
pathways_CC_terms <-read.csv('C:\\Users\\naata\\Documents\\MASTER\\BioSynth\\synthetic_env\\GSEA\\full_cc_terms.csv')
pathways_KEGG_terms <-read.csv('C:\\Users\\naata\\Documents\\MASTER\\TFM\\tfm\\KEGG_gs_descriptions.csv')

```

# OVERREPRESENTATION ANALYSIS

# Import results from DeSeq2 analysis
```{r}

DEG_HDAC<-read.table("C:\\Users\\naata\\Downloads\\Final_Condition_T_vs_C_LFC1_padj05.tsv", sep="\t", header = TRUE)
DEG_Stress <-read_csv("C:\\Users\\naata\\MASTER\\Enrichment\\Final_condition_PErep_vs_VMrep_padj0.05_woPErep3_LFC1filtered.csv")
DEG_HDAC_Stress<- read.table("C:\\Users\\naata\\MASTER\\Enrichment\\Condition_T_vs_VM_result_padj_0.05_DevStage.csv", sep= ",", header=TRUE)


DEG_ALL_HDAC <- DEG_HDAC$external_gene_name
DEG_ALL_Stress <- DEG_Stress$external_gene_name
DEG_ALL_HDAC_Stress <- DEG_HDAC_Stress$external_gene_name


DEG_UP_HDAC <- DEG_HDAC[which(DEG_HDAC$log2FoldChange > 1),]$external_gene_name
DEG_UP_Stress <- DEG_Stress[which(DEG_Stress$log2FoldChange > 1),]$external_gene_name
DEG_UP_HDAC_Stress <- DEG_HDAC_Stress[which(DEG_HDAC_Stress$log2FoldChange > 1),]$external_gene_name


DEG_DOWN_HDAC <- DEG_HDAC[which(DEG_HDAC$log2FoldChange < -1),]$external_gene_name
DEG_DOWN_Stress <- DEG_Stress[which(DEG_Stress$log2FoldChange < -1),]$external_gene_name
DEG_DOWN_HDAC_Stress <- DEG_HDAC_Stress[which(DEG_HDAC_Stress$log2FoldChange < -1),]$external_gene_name

```

# All genes
```{r}
colnames(pathways_BP_gs)<-c("bp", "gene")

# Get background genes : genes with detected minimal expression
colnames(background_ALL_gl_HDAC) = "gene"
bckgr_geneset_HDAC <- merge(background_ALL_gl_HDAC, pathways_BP_gs, by="gene")

colnames(background_ALL_gl_Stress) = "gene"
bckgr_geneset_Stress <- merge(background_ALL_gl_Stress, pathways_BP_gs, by="gene")

colnames(background_ALL_gl_HDAC_Stress) = "gene"
bckgr_geneset_HDAC_Stress <- merge(background_ALL_gl_HDAC_Stress, pathways_BP_gs, by="gene")


colnames(bckgr_geneset_HDAC)<-c("ID","bp")
bckgr_geneset_HDAC<-bckgr_geneset_HDAC[,c(2,1)]

colnames(bckgr_geneset_Stress)<-c("ID","bp")
bckgr_geneset_Stress<-bckgr_geneset_Stress[,c(2,1)]

colnames(bckgr_geneset_HDAC_Stress)<-c("ID","bp")
bckgr_geneset_HDAC_Stress<-bckgr_geneset_HDAC_Stress[,c(2,1)]

ORA_HDAC_ALL <- clusterProfiler::enricher(gene = DEG_ALL_HDAC, minGSSize=2, pvalueCutoff=0.05, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_HDAC, TERM2NAME = pathways_BP_terms)
ORA_Stress <- clusterProfiler::enricher(gene = DEG_ALL_Stress, minGSSize=2, pvalueCutoff=0.05, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_Stress, TERM2NAME = pathways_BP_terms)
ORA_HDAC_Stress <- clusterProfiler::enricher(gene = DEG_ALL_HDAC_Stress, minGSSize=2, pvalueCutoff=0.05, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_HDAC_Stress, TERM2NAME = pathways_BP_terms)


ORA_HDAC_ALL@organism="bnapus"
ORA_HDAC_ALL@ontology="BP"

ORA_Stress@organism="bnapus"
ORA_Stress@ontology="BP"

ORA_HDAC_Stress@organism="bnapus"
ORA_HDAC_Stress@ontology="BP"


HDAC_all <- importCP(ORA_HDAC_ALL, type = "other")
write_csv(HDAC_all, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\ORA_PE-SAHA_ALL_Results.csv")

Stress_all <- importCP(ORA_Stress, type = "other")
write_csv(Stress_all, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\ORA_HeatStress_ALL_Results.csv")

HDAC_Stress_all <- importCP(ORA_HDAC_Stress, type = "other")
write_csv(HDAC_Stress_all, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\ORA_HeatStress-SAHA_ALL_Results.csv")



##################### UP REGULATED GENES

colnames(background_UP_gl_HDAC) = "gene"
bckgr_geneset_HDAC <- merge(background_UP_gl_HDAC, pathways_BP_gs, by="gene")
colnames(bckgr_geneset_HDAC)<-c("bp","ID")
bckgr_geneset_HDAC<-bckgr_geneset_HDAC[,c(2,1)]


colnames(background_UP_gl_Stress)<-c("gene")
bckgr_geneset_Stress <- merge(background_UP_gl_Stress, pathways_BP_gs, by="gene")
colnames(bckgr_geneset_Stress)<-c("bp","ID")
bckgr_geneset_Stress<-bckgr_geneset_Stress[,c(2,1)]


colnames(background_UP_gl_HDAC_Stress) = "gene"
bckgr_geneset_HDAC_Stress <- merge(background_UP_gl_HDAC_Stress, pathways_BP_gs, by="gene")
colnames(bckgr_geneset_HDAC_Stress)<-c("bp","ID")
bckgr_geneset_HDAC_Stress<-bckgr_geneset_HDAC_Stress[,c(2,1)]



ORA_HDAC_UP <- clusterProfiler::enricher(gene = DEG_UP_HDAC, minGSSize=2, pvalueCutoff=0.05, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_HDAC, TERM2NAME = pathways_BP_terms)
ORA_Stress_UP <- clusterProfiler::enricher(gene = DEG_UP_Stress, minGSSize=2, pvalueCutoff=0.05, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_Stress, TERM2NAME = pathways_BP_terms)
ORA_HDAC_Stress_UP <- clusterProfiler::enricher(gene = DEG_UP_HDAC_Stress, minGSSize=2, pvalueCutoff=0.05, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_Stress, TERM2NAME = pathways_BP_terms)



ORA_HDAC_UP@organism="bnapus"
ORA_HDAC_UP@ontology="BP"

ORA_Stress_UP@organism="bnapus"
ORA_Stress_UP@ontology="BP"

ORA_HDAC_Stress_UP@organism="bnapus"
ORA_HDAC_Stress_UP@ontology="BP"



HDAC_up <- importCP(ORA_HDAC_UP, type = "other")
write_csv(HDAC_up, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\ORA_PE_SAHA_up_Results.csv")

Stress_up <- importCP(ORA_Stress_UP, type = "other")
write_csv(Stress_up, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\ORA_HeatStress_up_Results.csv")

HDAC_Stress_up <- importCP(ORA_HDAC_Stress_UP, type = "other")
write_csv(HDAC_Stress_up, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\ORA_HeatStress_SAHA_up_Results.csv")



############################### DOWNREGULATED GENES

colnames(background_DOWN_gl_HDAC) = "gene"
bckgr_geneset_HDAC <- merge(background_DOWN_gl_HDAC, pathways_BP_gs, by="gene")

colnames(background_DOWN_gl_Stress) = "gene"
bckgr_geneset_Stress <- merge(background_DOWN_gl_Stress, pathways_BP_gs, by="gene")

colnames(background_DOWN_gl_HDAC_Stress) = "gene"
bckgr_geneset_HDAC_Stress <- merge(background_DOWN_gl_HDAC_Stress, pathways_BP_gs, by="gene")


colnames(bckgr_geneset_HDAC)<-c("ID","bp")
bckgr_geneset_HDAC<-bckgr_geneset_HDAC[,c(2,1)]

colnames(bckgr_geneset_Stress)<-c("ID","bp")
bckgr_geneset_Stress<-bckgr_geneset_Stress[,c(2,1)]

colnames(bckgr_geneset_HDAC_Stress)<-c("ID","bp")
bckgr_geneset_HDAC_Stress<-bckgr_geneset_HDAC_Stress[,c(2,1)]



ORA_HDAC_DOWN <- clusterProfiler::enricher(gene = DEG_DOWN_HDAC, minGSSize=2, pvalueCutoff=0.05, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_HDAC, TERM2NAME = pathways_BP_terms)
ORA_Stress_DOWN <- clusterProfiler::enricher(gene = DEG_DOWN_Stress, minGSSize=2, pvalueCutoff=0.05, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_Stress, TERM2NAME = pathways_BP_terms)
ORA_HDAC_Stress_DOWN <- clusterProfiler::enricher(gene = DEG_DOWN_HDAC_Stress, minGSSize=2, pvalueCutoff=0.05, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_Stress, TERM2NAME = pathways_BP_terms)


ORA_HDAC_DOWN@organism="bnapus"
ORA_HDAC_DOWN@ontology="BP"

ORA_Stress_DOWN@organism="bnapus"
ORA_Stress_DOWN@ontology="BP"

ORA_HDAC_Stress_DOWN@organism="bnapus"
ORA_HDAC_Stress_DOWN@ontology="BP"



HDAC_down <- importCP(ORA_HDAC_DOWN, type = "other")
write_csv(HDAC_down, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\ORA_PE_SAHA_down_Results.csv")

Stress_down <- importCP(ORA_Stress_DOWN, type = "other")
write_csv(Stress_down, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\ORA_HeatStress_down_Results.csv")

HDAC_Stress_down <- importCP(ORA_HDAC_Stress_DOWN, type = "other")
write_csv(HDAC_Stress_down, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\ORA_HeatStress_SAHA_down_Results.csv")









```

```{r}
# Calculate similarity matrix
simMatrix_HDACi <- calculateSimMatrix(HDAC_up$ID,
                                orgdb=org.Brassicanapus.eg.db,
                                ont="BP",
                                method="Wang")
# Reducing terms
scores <- setNames(-log10(HDAC_up$qvalue), HDAC_up$ID)
reducedTerms_HDAC <- reduceSimMatrix(simMatrix_HDACi,
                                scores,
                                threshold=0.9,
                                orgdb=org.Brassicanapus.eg.db)

HDAC_up_plot = HDAC_up[HDAC_up$ID %in% reducedTerms_HDAC$parent,]
HDAC_up_plot_2 = HDAC_up[HDAC_up$ID %in% reducedTerms_HDAC$parent,]



plotEnrichAdv(HDAC_up_plot, HDAC_down, plot_type = "one") + theme(panel.background = element_rect(fill="#ECFFDC"))

HDAC_up_plot$Description =  gsub("regulation", "reg.", HDAC_up_plot$Description)
HDAC_up_plot$Description =  gsub("process", "proc.", HDAC_up_plot$Description)
HDAC_up_plot$Description =  gsub("metabolic process", "met.proc.", HDAC_up_plot$Description)
HDAC_up_plot$Description =  gsub("catabolic process", "cat.proc.", HDAC_up_plot$Description)
HDAC_up_plot$Description =  gsub("phosphoenolpyruvate", "PEP", HDAC_up_plot$Description)
HDAC_up_plot$Description =  gsub("between organisms", "", HDAC_up_plot$Description)
HDAC_up_plot$Description =  gsub("between organisms", "", HDAC_up_plot$Description)
HDAC_up_plot$Description = gsub("Erythrose", "\n\nErythrose", HDAC_up_plot$Description)
HDAC_up_plot$Description = gsub("Biological Process", "BP", HDAC_up_plot$Description)
HDAC_down$Description = gsub("non-coding RNA", "ncRNA", HDAC_down$Description)
HDAC_down$Description = gsub("DNA methylation", "DNA methyl.", HDAC_down$Description)
HDAC_down$Description = gsub("organization", "org.", HDAC_down$Description)
HDAC_down$Description = gsub("formation", "form.", HDAC_down$Description)
HDAC_down$Description = gsub("epigenetic", "epig.", HDAC_down$Description)




svg("C:\\Users\\naata\\Downloads\\GO_BP_up_down_f20.svg", width=6.15, height=4)
plotEnrichAdv(HDAC_up_plot[1:20,], HDAC_down, stats_metric="p.adjust", term_metric = 'FoldEnrichment', plot_type = "one", xlim_left=-16, xlim_right = 13, color=c("#00A36C","coral")) + labs(caption = "GO Biological Process") + theme(panel.background = element_rect(fill="#ECFFDC"), panel.border = element_blank())
dev.off()


svg("C:\\Users\\naata\\Downloads\\GO_BP_up_down_two.svg", width=15, height=6)
plotEnrichAdv(HDAC_up_plot, HDAC_down, stats_metric="p.adjust", term_metric = 'Count', plot_type = "two", xlim_left=-16, xlim_right = 13, color=c("coral", "grey","#00A36C", "grey")) + labs(title = "GO Biological Process") + theme(panel.background = element_rect(fill="#ECFFDC"), panel.border = element_blank(), text=element_text(color="green"))
dev.off()



svg("C:\\Users\\naata\\Downloads\\GO_CC_up_down.svg", width=5.55, height=3)
plotEnrichAdv(HDAC_CC_UP, HDAC_CC_DOWN, stats_metric="p.adjust", term_metric = 'FoldEnrichment', plot_type = "one", xlim_left=-16.1, xlim_right = 15, color=c("#00A36C","coral")) + labs(caption = "GO Cellular Component") + theme(panel.background = element_rect(fill="#FCDCC8"), panel.border = element_blank())
dev.off()





HDAC_MF_UP$Description = gsub("activity", "act.", HDAC_MF_UP$Description)
HDAC_MF_UP$Description = gsub("reduction", "reduct", HDAC_MF_UP$Description)
HDAC_MF_UP$Description = gsub("incorporation", "inc.", HDAC_MF_UP$Description)
HDAC_MF_UP$Description = gsub("oxygen", "O", HDAC_MF_UP$Description)
HDAC_MF_UP$Description = gsub("donor", "d", HDAC_MF_UP$Description)
HDAC_MF_UP$Description = gsub("acceptor", "acc", HDAC_MF_UP$Description)
HDAC_MF_UP$Description = gsub("transferase act.,", "", HDAC_MF_UP$Description)
HDAC_MF_UP$Description = gsub("Oxidorreductase", "Oxidorred.", HDAC_MF_UP$Description)
HDAC_MF_UP = HDAC_MF_UP[!(grepl("acting on", HDAC_MF_UP$Description)),]

HDAC_MF_DOWN$Description = gsub("activity", "act.", HDAC_MF_DOWN$Description)



svg("C:\\Users\\naata\\Downloads\\GO_MF_up_down_f20.svg", width=6.15, height=4)
plotEnrichAdv(HDAC_MF_UP[1:20,], HDAC_MF_DOWN, stats_metric="p.adjust", term_metric = 'FoldEnrichment', plot_type = "one", xlim_left=-18.1, xlim_right = 18, color=c("#00A36C","coral")) + labs(caption = "GO Molecular Function") + theme(panel.background = element_rect(fill="#FFFDE3"), panel.border = element_blank())
dev.off()

# svg("C:\\Users\\naata\\Downloads\\GO_KEGG_up_down_f20.svg", width=6.15, height=7)
# plotEnrich(HDAC_KEGG_UP[1:20,], stats_metric="p.adjust", term_metric = 'Count', plot_type = "bar", xlim_left=0, xlim_right = 20) + labs(caption = "KEGG") + theme(panel.background = element_rect(fill="#FFFDE3"), panel.border = element_blank())
# dev.off()

svg("C:\\Users\\naata\\Downloads\\GO_KEGG_up_down.svg", width=5.15, height=1.75)
ggplot(HDAC_KEGG_UP, aes(x = -log10(p.adjust), y = reorder(Description,-log10(p.adjust)))) + 
  geom_bar(stat="identity",fill="coral", width=0.7) +
  theme_bw(base_size = 14) +
  ylab(NULL) +
  labs(caption = "KEGG pathways") +theme_minimal()+theme(panel.grid.minor=element_blank(), 
                                                         panel.grid.major=element_blank(), 
                                                         axis.text.y=element_text(colour="black"),
                                                         axis.title.y=element_text(size=10),
                                                         panel.background = element_rect(fill="#bae1ff",linewidth = 0),
                                                         plot.background = element_rect(fill="#bae1ff",linewidth = 0),
                                                         panel.border = element_blank())
dev.off()







HDAC_Stress_all[HDAC_Stress_all$ID %in% Stress_all$ID,]




svg("C:\\Users\\naata\\Downloads\\GO_MF_up_down_f20.svg", width=6.15, height=4)
plotEnrichAdv(HDAC_MF_UP[1:20,], HDAC_MF_DOWN, stats_metric="p.adjust", term_metric = 'FoldEnrichment', plot_type = "one", xlim_left=-18.1, xlim_right = 18, color=c("#00A36C","coral")) + labs(caption = "GO Molecular Function") + theme(panel.background = element_rect(fill="#FFFDE3"), panel.border = element_blank())
dev.off()






######################## ME COMPARISON



# Calculate similarity matrix
simMatrix_HDACi <- calculateSimMatrix(HDAC_Stress_all$ID,
                                orgdb=org.Brassicanapus.eg.db,
                                ont="BP",
                                method="Wang")
# Reducing terms
scores <- setNames(-log10(HDAC_Stress_all$qvalue), HDAC_Stress_all$ID)
reducedTerms_HDAC <- reduceSimMatrix(simMatrix_HDACi,
                                scores,
                                threshold=0.7,
                                orgdb=org.Brassicanapus.eg.db)

HSSAHA_toplot = HDAC_Stress_all[HDAC_Stress_all$ID %in% reducedTerms_HDAC$parent,]

p_BP_SAHA = plotEnrich(HSSAHA_toplot, plot_type="lollipop", up_color = "coral", down_color ="grey", scale_ratio = 0.03 ) + labs(title = "HeatStress-SAHA") + theme(panel.background = element_rect(fill="#ECFFDC"), panel.border = element_blank())

# Calculate similarity matrix
simMatrix_HDACi <- calculateSimMatrix(Stress_all$ID,
                                orgdb=org.Brassicanapus.eg.db,
                                ont="BP",
                                method="Wang")
# Reducing terms
scores <- setNames(-log10(Stress_all$qvalue), Stress_all$ID)
reducedTerms_HDAC <- reduceSimMatrix(simMatrix_HDACi,
                                scores,
                                threshold=0.7,
                                orgdb=org.Brassicanapus.eg.db)

HS_toplot = Stress_all[Stress_all$ID %in% reducedTerms_HDAC$parent,]

p_BP_Stress = plotEnrich(HS_toplot , plot_type="lollipop", up_color = "coral", down_color ="grey", scale_ratio = 0.03) + labs(title = "HeatStress") + theme(panel.background = element_rect(fill="#ECFFDC"), panel.border = element_blank())


row_BP_comp = ggarrange(p_BP_SAHA, p_BP_Stress, ncol=2, widths=c(5.5, 5))
ggsave("C:\\Users\\naata\\Downloads\\ME_comparison_BP.svg", row_BP_comp, width=11, height=5)





# Calculate similarity matrix
simMatrix_HDACi <- calculateSimMatrix(HDAC_Stress_CC_all$ID,
                                orgdb=org.Brassicanapus.eg.db,
                                ont="CC",
                                method="Wang")
# Reducing terms
scores <- setNames(-log10(HDAC_Stress_CC_all$qvalue), HDAC_Stress_CC_all$ID)
reducedTerms_HDAC <- reduceSimMatrix(simMatrix_HDACi,
                                scores,
                                threshold=0.4,
                                orgdb=org.Brassicanapus.eg.db)

HSSAHA_toplot_CC = HDAC_Stress_CC_all[HDAC_Stress_CC_all$ID %in% reducedTerms_HDAC$parent,]

p_CC_SAHA = plotEnrich(HSSAHA_toplot_CC, plot_type="lollipop", up_color = "coral", down_color ="grey", scale_ratio = 0.03 ) + theme(panel.background = element_rect(fill="#FCDCC8"), panel.border = element_blank())

# Calculate similarity matrix
simMatrix_HDACi <- calculateSimMatrix(Stress_CC_all$ID,
                                orgdb=org.Brassicanapus.eg.db,
                                ont="CC",
                                method="Wang")
# Reducing terms
scores <- setNames(-log10(Stress_CC_all$qvalue), Stress_CC_all$ID)
reducedTerms_HDAC <- reduceSimMatrix(simMatrix_HDACi,
                                scores,
                                threshold=0.4,
                                orgdb=org.Brassicanapus.eg.db)

HS_toplot_CC = Stress_CC_all[Stress_CC_all$ID %in% reducedTerms_HDAC$parent,]

p_CC_Stress = plotEnrich(HS_toplot_CC , plot_type="lollipop", up_color = "coral", down_color ="grey", scale_ratio = 0.03) + theme(panel.background = element_rect(fill="#FCDCC8"), panel.border = element_blank())


row_CC_comp = ggarrange(p_CC_SAHA, p_CC_Stress, ncol=2, widths=c(5.5, 5))
ggsave("C:\\Users\\naata\\Downloads\\ME_comparison_CC.svg", row_CC_comp, width=11, height=4.2)




# Calculate similarity matrix
simMatrix_HDACi <- calculateSimMatrix(HDAC_Stress_MF_all$ID,
                                orgdb=org.Brassicanapus.eg.db,
                                ont="MF",
                                method="Wang")
# Reducing terms
scores <- setNames(-log10(HDAC_Stress_MF_all$qvalue), HDAC_Stress_MF_all$ID)
reducedTerms_HDAC <- reduceSimMatrix(simMatrix_HDACi,
                                scores,
                                threshold=0.8,
                                orgdb=org.Brassicanapus.eg.db)

HSSAHA_toplot_MF = HDAC_Stress_MF_all[HDAC_Stress_MF_all$ID %in% reducedTerms_HDAC$parent,]

p_MF_SAHA = plotEnrich(HSSAHA_toplot_MF, plot_type="lollipop", up_color = "coral", down_color ="grey", scale_ratio = 0.03 ) + theme(panel.background = element_rect(fill="#FFFDE3"), panel.border = element_blank())

# Calculate similarity matrix
simMatrix_HDACi <- calculateSimMatrix(Stress_MF_all$ID,
                                orgdb=org.Brassicanapus.eg.db,
                                ont="MF",
                                method="Wang")
# Reducing terms
scores <- setNames(-log10(Stress_MF_all$qvalue), Stress_MF_all$ID)
reducedTerms_HDAC <- reduceSimMatrix(simMatrix_HDACi,
                                scores,
                                threshold=0.8,
                                orgdb=org.Brassicanapus.eg.db)

HS_toplot_MF = Stress_MF_all[Stress_MF_all$ID %in% reducedTerms_HDAC$parent,]

p_MF_Stress = plotEnrich(HS_toplot_MF , plot_type="lollipop", up_color = "coral", down_color ="grey", scale_ratio = 0.03) + theme(panel.background = element_rect(fill="#FFFDE3"), panel.border = element_blank())


row_MF_comp = ggarrange(p_MF_SAHA, p_MF_Stress, ncol=2, widths=c(3.5, 5))
ggsave("C:\\Users\\naata\\Downloads\\ME_comparison_MF.svg", row_MF_comp, width=11, height=3.1)



p_KEGG_S <- plotEnrichAdv(HDAC_Stress_KEGG_UP, HDAC_Stress_KEGG_DOWN[2:10,], stats_metric="p.adjust", term_metric = 'FoldEnrichment', plot_type = "one", xlim_left=-12.5, xlim_right = 12, color=c("#00A36C","coral")) + labs(caption = "KEGG") + theme(panel.background = element_rect(fill="#bae1ff"), panel.border = element_blank())


p_KEGG_Stress <- plotEnrichAdv(Stress_KEGG_UP, Stress_KEGG_DOWN[2:5,], stats_metric="p.adjust", term_metric = 'FoldEnrichment', plot_type = "one", xlim_left=-11.2, xlim_right = 11, color=c("#00A36C","coral")) + labs(caption = "KEGG") + theme(panel.background = element_rect(fill="#bae1ff"), panel.border = element_blank())


row_KEGG_comp = ggarrange(p_KEGG_S, p_KEGG_Stress, ncol=2, widths=c(5, 5), common.legend = TRUE, legend = "bottom")
ggsave("C:\\Users\\naata\\Downloads\\ME_comparison_KEGG_2.svg", row_KEGG_comp, width=11, height=2.5)




```



# CC

# All genes
```{r}
colnames(pathways_CC_gs)<-c("cc", "gene")

# Get background genes : genes with detected minimal expression

bckgr_geneset_HDAC <- merge(background_ALL_gl_HDAC, pathways_CC_gs, by="gene")
bckgr_geneset_Stress <- merge(background_ALL_gl_Stress, pathways_CC_gs, by="gene")
bckgr_geneset_HDAC_Stress <- merge(background_ALL_gl_HDAC_Stress, pathways_CC_gs, by="gene")

colnames(bckgr_geneset_HDAC)<-c("ID","cc")
bckgr_geneset_HDAC<-bckgr_geneset_HDAC[,c(2,1)]

colnames(bckgr_geneset_Stress)<-c("ID","cc")
bckgr_geneset_Stress<-bckgr_geneset_Stress[,c(2,1)]

colnames(bckgr_geneset_HDAC_Stress)<-c("ID","cc")
bckgr_geneset_HDAC_Stress<-bckgr_geneset_HDAC_Stress[,c(2,1)]



ORA_HDAC_CC <- clusterProfiler::enricher(gene = DEG_ALL_HDAC, minGSSize=2, pvalueCutoff=1, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_HDAC, TERM2NAME = pathways_CC_terms)
ORA_Stress_CC <- clusterProfiler::enricher(gene = DEG_ALL_Stress, minGSSize=2, pvalueCutoff=0.05, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_Stress, TERM2NAME = pathways_CC_terms)
ORA_HDAC_Stress_CC <- clusterProfiler::enricher(gene = DEG_ALL_HDAC_Stress, minGSSize=2, pvalueCutoff=0.05, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_HDAC_Stress, TERM2NAME = pathways_CC_terms)



ORA_HDAC_CC@organism="bnapus"
ORA_HDAC_CC@ontology="CC"

ORA_Stress_CC@organism="bnapus"
ORA_Stress_CC@ontology="CC"

ORA_HDAC_Stress_CC@organism="bnapus"
ORA_HDAC_Stress_CC@ontology="CC"



HDAC_CC_all <- importCP(ORA_HDAC_CC, type = "other")
write_csv(HDAC_CC_all, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\ORA_HDAC_ALL_CC_Results.csv")

Stress_CC_all <- importCP(ORA_Stress_CC, type = "other")
write_csv(Stress_CC_all, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\ORA_HeatStress_ALL_CC_Results.csv")

HDAC_Stress_CC_all <- importCP(ORA_HDAC_Stress_CC, type = "other")
write_csv(HDAC_Stress_CC_all, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\ORA_HeatStress_SAHA_ALL_CC_Results.csv")



################## UP REGULATED GENES

colnames(background_UP_gl_HDAC) = "gene"
bckgr_geneset_HDAC <- merge(background_UP_gl_HDAC, pathways_CC_gs, by="gene")

colnames(background_UP_gl_Stress) = "gene"
bckgr_geneset_Stress <- merge(background_UP_gl_Stress, pathways_CC_gs, by="gene")

colnames(background_UP_gl_HDAC_Stress) = "gene"
bckgr_geneset_HDAC_Stress <- merge(background_UP_gl_HDAC_Stress, pathways_CC_gs, by="gene")


colnames(bckgr_geneset_HDAC)<-c("ID","cc")
bckgr_geneset_HDAC<-bckgr_geneset_HDAC[,c(2,1)]

colnames(bckgr_geneset_Stress)<-c("ID","cc")
bckgr_geneset_Stress<-bckgr_geneset_Stress[,c(2,1)]

colnames(bckgr_geneset_HDAC_Stress)<-c("ID","cc")
bckgr_geneset_HDAC_Stress<-bckgr_geneset_HDAC_Stress[,c(2,1)]


ORA_HDAC_CC_up <- clusterProfiler::enricher(gene = DEG_UP_HDAC, minGSSize=2, pvalueCutoff=0.05, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_HDAC, TERM2NAME = pathways_CC_terms)
ORA_Stress_CC_up <- clusterProfiler::enricher(gene = DEG_UP_Stress, minGSSize=2, pvalueCutoff=0.05, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_Stress, TERM2NAME = pathways_CC_terms)
ORA_HDAC_Stress_CC_up <- clusterProfiler::enricher(gene = DEG_UP_HDAC_Stress, minGSSize=2, pvalueCutoff=0.05, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_HDAC_Stress, TERM2NAME = pathways_CC_terms)


ORA_HDAC_CC_up@organism="bnapus"
ORA_HDAC_CC_up@ontology="CC"

ORA_Stress_CC_up@organism="bnapus"
ORA_Stress_CC_up@ontology="CC"

ORA_HDAC_Stress_CC_up@organism="bnapus"
ORA_HDAC_Stress_CC_up@ontology="CC"



HDAC_CC_UP <- importCP(ORA_HDAC_CC_up, type = "other")
write_csv(HDAC_CC_UP, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\ORA_PE_SAHA_UP_CC_Results.csv")

Stress_CC_UP <- importCP(ORA_Stress_CC_up, type = "other")
write_csv(Stress_CC_UP, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\ORA_HeatStress_UP_CC_Results.csv")

HDAC_Stress_CC_UP <- importCP(ORA_HDAC_Stress_CC_up, type = "other")
write_csv(HDAC_Stress_CC_UP, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\ORA_HeatStress_SAHA_UP_CC_Results.csv")


################# DOWN REGULATED GENES


colnames(background_DOWN_gl_HDAC) = "gene"
bckgr_geneset_HDAC <- merge(background_DOWN_gl_HDAC, pathways_CC_gs, by="gene")

colnames(background_DOWN_gl_Stress) = "gene"
bckgr_geneset_Stress <- merge(background_DOWN_gl_Stress, pathways_CC_gs, by="gene")

colnames(background_DOWN_gl_HDAC_Stress) = "gene"
bckgr_geneset_HDAC_Stress <- merge(background_DOWN_gl_HDAC_Stress, pathways_CC_gs, by="gene")


colnames(bckgr_geneset_HDAC)<-c("ID","cc")
bckgr_geneset_HDAC<-bckgr_geneset_HDAC[,c(2,1)]

colnames(bckgr_geneset_Stress)<-c("ID","cc")
bckgr_geneset_Stress<-bckgr_geneset_Stress[,c(2,1)]

colnames(bckgr_geneset_HDAC_Stress)<-c("ID","cc")
bckgr_geneset_HDAC_Stress<-bckgr_geneset_HDAC_Stress[,c(2,1)]



ORA_HDAC_CC_down <- clusterProfiler::enricher(gene = DEG_DOWN_HDAC, minGSSize=2, pvalueCutoff=0.05, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_HDAC, TERM2NAME = pathways_CC_terms)
ORA_Stress_CC_down <- clusterProfiler::enricher(gene = DEG_DOWN_Stress, minGSSize=2, pvalueCutoff=0.05, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_Stress, TERM2NAME = pathways_CC_terms)
ORA_HDAC_Stress_CC_down <- clusterProfiler::enricher(gene = DEG_DOWN_HDAC_Stress, minGSSize=2, pvalueCutoff=0.05, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_HDAC_Stress, TERM2NAME = pathways_CC_terms)


ORA_HDAC_CC_down@organism="bnapus"
ORA_HDAC_CC_down@ontology="CC"

ORA_Stress_CC_down@organism="bnapus"
ORA_Stress_CC_down@ontology="CC"

ORA_HDAC_Stress_CC_down@organism="bnapus"
ORA_HDAC_Stress_CC_down@ontology="CC"


HDAC_CC_DOWN <- importCP(ORA_HDAC_CC_down, type = "other")
write_csv(HDAC_CC_DOWN, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\ORA_PE_SAHA_DOWN_CC_Results.csv")

Stress_CC_DOWN <- importCP(ORA_Stress_CC_down, type = "other")
write_csv(Stress_CC_DOWN, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\ORA_HeatStress_DOWN_CC_Results.csv")

HDAC_Stress_CC_DOWN <- importCP(ORA_HDAC_Stress_CC_down, type = "other")
write_csv(HDAC_Stress_CC_all, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\ORA_HeatStress_SAHA_DOWN_CC_Results.csv")



```

# MF

# All genes
```{r}
colnames(pathways_MF_gs)<-c("mf", "gene")

# Get background genes : genes with detected minimal expression

bckgr_geneset_HDAC <- merge(background_ALL_gl_HDAC, pathways_MF_gs, by="gene")
bckgr_geneset_Stress <- merge(background_ALL_gl_Stress, pathways_MF_gs, by="gene")
bckgr_geneset_HDAC_Stress <- merge(background_ALL_gl_HDAC_Stress, pathways_MF_gs, by="gene")

colnames(bckgr_geneset_HDAC)<-c("ID","mf")
bckgr_geneset_HDAC<-bckgr_geneset_HDAC[,c(2,1)]

colnames(bckgr_geneset_Stress)<-c("ID","mf")
bckgr_geneset_Stress<-bckgr_geneset_Stress[,c(2,1)]

colnames(bckgr_geneset_HDAC_Stress)<-c("ID","mf")
bckgr_geneset_HDAC_Stress<-bckgr_geneset_HDAC_Stress[,c(2,1)]



ORA_HDAC_MF <- clusterProfiler::enricher(gene = DEG_ALL_HDAC, minGSSize=2, pvalueCutoff=1, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_HDAC, TERM2NAME = pathways_MF_terms)
ORA_Stress_MF <- clusterProfiler::enricher(gene = DEG_ALL_Stress, minGSSize=2, pvalueCutoff=0.05, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_Stress, TERM2NAME = pathways_MF_terms)
ORA_HDAC_Stress_MF <- clusterProfiler::enricher(gene = DEG_ALL_HDAC_Stress, minGSSize=2, pvalueCutoff=0.05, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_HDAC_Stress, TERM2NAME = pathways_MF_terms)



ORA_HDAC_MF@organism="bnapus"
ORA_HDAC_MF@ontology="MF"

ORA_Stress_MF@organism="bnapus"
ORA_Stress_MF@ontology="MF"

ORA_HDAC_Stress_MF@organism="bnapus"
ORA_HDAC_Stress_MF@ontology="MF"



HDAC_MF_all <- importCP(ORA_HDAC_MF, type = "other")
write_csv(HDAC_MF_all, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\ORA_HDAC_ALL_MF_Results.csv")

Stress_MF_all <- importCP(ORA_Stress_MF, type = "other")
write_csv(Stress_MF_all, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\ORA_HeatStress_ALL_MF_Results.csv")

HDAC_Stress_MF_all <- importCP(ORA_HDAC_Stress_MF, type = "other")
write_csv(HDAC_Stress_MF_all, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\ORA_HeatStress_SAHA_ALL_MF_Results.csv")



################## UP REGULATED GENES

colnames(background_UP_gl_HDAC) = "gene"
bckgr_geneset_HDAC <- merge(background_UP_gl_HDAC, pathways_MF_gs, by="gene")

colnames(background_UP_gl_Stress) = "gene"
bckgr_geneset_Stress <- merge(background_UP_gl_Stress, pathways_MF_gs, by="gene")

colnames(background_UP_gl_HDAC_Stress) = "gene"
bckgr_geneset_HDAC_Stress <- merge(background_UP_gl_HDAC_Stress, pathways_MF_gs, by="gene")


colnames(bckgr_geneset_HDAC)<-c("ID","mf")
bckgr_geneset_HDAC<-bckgr_geneset_HDAC[,c(2,1)]

colnames(bckgr_geneset_Stress)<-c("ID","mf")
bckgr_geneset_Stress<-bckgr_geneset_Stress[,c(2,1)]

colnames(bckgr_geneset_HDAC_Stress)<-c("ID","mf")
bckgr_geneset_HDAC_Stress<-bckgr_geneset_HDAC_Stress[,c(2,1)]


ORA_HDAC_MF_up <- clusterProfiler::enricher(gene = DEG_UP_HDAC, minGSSize=2, pvalueCutoff=0.05, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_HDAC, TERM2NAME = pathways_MF_terms)
ORA_Stress_MF_up <- clusterProfiler::enricher(gene = DEG_UP_Stress, minGSSize=2, pvalueCutoff=0.05, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_Stress, TERM2NAME = pathways_MF_terms)
ORA_HDAC_Stress_MF_up <- clusterProfiler::enricher(gene = DEG_UP_HDAC_Stress, minGSSize=2, pvalueCutoff=0.05, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_HDAC_Stress, TERM2NAME = pathways_MF_terms)


ORA_HDAC_MF_up@organism="bnapus"
ORA_HDAC_MF_up@ontology="MF"

ORA_Stress_MF_up@organism="bnapus"
ORA_Stress_MF_up@ontology="MF"

ORA_HDAC_Stress_MF_up@organism="bnapus"
ORA_HDAC_Stress_MF_up@ontology="MF"



HDAC_MF_UP <- importCP(ORA_HDAC_MF_up, type = "other")
write_csv(HDAC_MF_UP, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\ORA_PE_SAHA_UP_MF_Results.csv")

Stress_MF_UP <- importCP(ORA_Stress_MF_up, type = "other")
write_csv(Stress_MF_UP, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\ORA_HeatStress_UP_MF_Results.csv")

HDAC_Stress_MF_UP <- importCP(ORA_HDAC_Stress_MF_up, type = "other")
write_csv(HDAC_Stress_MF_UP, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\ORA_HeatStress_SAHA_UP_MF_Results.csv")


################# DOWN REGULATED GENES


colnames(background_DOWN_gl_HDAC) = "gene"
bckgr_geneset_HDAC <- merge(background_DOWN_gl_HDAC, pathways_MF_gs, by="gene")

colnames(background_DOWN_gl_Stress) = "gene"
bckgr_geneset_Stress <- merge(background_DOWN_gl_Stress, pathways_MF_gs, by="gene")

colnames(background_DOWN_gl_HDAC_Stress) = "gene"
bckgr_geneset_HDAC_Stress <- merge(background_DOWN_gl_HDAC_Stress, pathways_MF_gs, by="gene")


colnames(bckgr_geneset_HDAC)<-c("ID","mf")
bckgr_geneset_HDAC<-bckgr_geneset_HDAC[,c(2,1)]

colnames(bckgr_geneset_Stress)<-c("ID","mf")
bckgr_geneset_Stress<-bckgr_geneset_Stress[,c(2,1)]

colnames(bckgr_geneset_Stress)<-c("ID","mf")
bckgr_geneset_HDAC_Stress<-bckgr_geneset_HDAC_Stress[,c(2,1)]



ORA_HDAC_MF_down <- clusterProfiler::enricher(gene = DEG_DOWN_HDAC, minGSSize=2, pvalueCutoff=0.05, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_HDAC, TERM2NAME = pathways_MF_terms)
ORA_Stress_MF_down <- clusterProfiler::enricher(gene = DEG_DOWN_Stress, minGSSize=2, pvalueCutoff=0.05, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_Stress, TERM2NAME = pathways_MF_terms)
ORA_HDAC_Stress_MF_down <- clusterProfiler::enricher(gene = DEG_DOWN_HDAC_Stress, minGSSize=2, pvalueCutoff=0.05, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_HDAC_Stress, TERM2NAME = pathways_MF_terms)


ORA_HDAC_MF_down@organism="bnapus"
ORA_HDAC_MF_down@ontology="MF"

ORA_Stress_MF_down@organism="bnapus"
ORA_Stress_MF_down@ontology="MF"

ORA_HDAC_Stress_MF_down@organism="bnapus"
ORA_HDAC_Stress_MF_down@ontology="MF"


HDAC_MF_DOWN <- importCP(ORA_HDAC_MF_down, type = "other")
write_csv(HDAC_MF_DOWN, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\ORA_PE_SAHA_DOWN_MF_Results.csv")

Stress_MF_DOWN <- importCP(ORA_Stress_MF_down, type = "other")
write_csv(Stress_MF_DOWN, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\ORA_HeatStress_DOWN_MF_Results.csv")

HDAC_Stress_MF_DOWN <- importCP(ORA_HDAC_Stress_MF_down, type = "other")
write_csv(HDAC_Stress_MF_all, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\ORA_HeatStress_SAHA_DOWN_MF_Results.csv")


```

# KEGG

# All genes
```{r}
colnames(pathways_KEGG_gs)<-c("bp", "gene")

# Get background genes : genes with detected minimal expression

bckgr_geneset_HDAC <- merge(background_ALL_gl_HDAC, pathways_KEGG_gs, by="gene")
bckgr_geneset_Stress <- merge(background_ALL_gl_Stress, pathways_KEGG_gs, by="gene")
bckgr_geneset_HDAC_Stress <- merge(background_ALL_gl_HDAC_Stress, pathways_KEGG_gs, by="gene")

colnames(bckgr_geneset_HDAC)<-c("ID","bp")
bckgr_geneset_HDAC<-bckgr_geneset_HDAC[,c(2,1)]

colnames(bckgr_geneset_Stress)<-c("ID","bp")
bckgr_geneset_Stress<-bckgr_geneset_Stress[,c(2,1)]

colnames(bckgr_geneset_HDAC_Stress)<-c("ID","bp")
bckgr_geneset_HDAC_Stress<-bckgr_geneset_HDAC_Stress[,c(2,1)]



ORA_HDAC_KEGG <- clusterProfiler::enricher(gene = DEG_ALL_HDAC, minGSSize=2, pvalueCutoff=1, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_HDAC, TERM2NAME = pathways_KEGG_terms)
ORA_Stress_KEGG <- clusterProfiler::enricher(gene = DEG_ALL_Stress, minGSSize=2, pvalueCutoff=0.05, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_Stress, TERM2NAME = pathways_KEGG_terms)
ORA_HDAC_Stress_KEGG <- clusterProfiler::enricher(gene = DEG_ALL_HDAC_Stress, minGSSize=2, pvalueCutoff=0.05, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_HDAC_Stress, TERM2NAME = pathways_KEGG_terms)



ORA_HDAC_KEGG@organism="bnapus"
ORA_HDAC_KEGG@ontology="BP"

ORA_Stress_KEGG@organism="bnapus"
ORA_Stress_KEGG@ontology="BP"

ORA_HDAC_Stress_KEGG@organism="bnapus"
ORA_HDAC_Stress_KEGG@ontology="BP"



HDAC_KEGG_all <- importCP(ORA_HDAC_KEGG, type = "other")
write_csv(HDAC_KEGG_all, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\ORA_HDAC_ALL_KEGG_Results.csv")

Stress_KEGG_all <- importCP(ORA_Stress_KEGG, type = "other")
write_csv(Stress_KEGG_all, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\ORA_HeatStress_ALL_KEGG_Results.csv")

HDAC_Stress_KEGG_all <- importCP(ORA_HDAC_Stress_KEGG, type = "other")
write_csv(HDAC_Stress_KEGG_all, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\ORA_HeatStress_SAHA_ALL_KEGG_Results.csv")



################## UP REGULATED GENES

bckgr_geneset_HDAC <- merge(background_UP_gl_HDAC, pathways_KEGG_gs, by="gene")
bckgr_geneset_Stress <- merge(background_UP_gl_Stress, pathways_KEGG_gs, by="gene")
bckgr_geneset_HDAC_Stress <- merge(background_UP_gl_HDAC_Stress, pathways_KEGG_gs, by="gene")

colnames(bckgr_geneset_HDAC)<-c("ID","bp")
bckgr_geneset_HDAC<-bckgr_geneset_HDAC[,c(2,1)]

colnames(bckgr_geneset_Stress)<-c("ID","bp")
bckgr_geneset_Stress<-bckgr_geneset_Stress[,c(2,1)]

colnames(bckgr_geneset_HDAC_Stress)<-c("ID","bp")
bckgr_geneset_HDAC_Stress<-bckgr_geneset_HDAC_Stress[,c(2,1)]



ORA_HDAC_KEGG_up <- clusterProfiler::enricher(gene = DEG_UP_HDAC, minGSSize=2, pvalueCutoff=0.05, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_HDAC, TERM2NAME = pathways_KEGG_terms)
ORA_Stress_KEGG_up <- clusterProfiler::enricher(gene = DEG_UP_Stress, minGSSize=2, pvalueCutoff=0.05, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_Stress, TERM2NAME = pathways_KEGG_terms)
ORA_HDAC_Stress_KEGG_up <- clusterProfiler::enricher(gene = DEG_UP_HDAC_Stress, minGSSize=2, pvalueCutoff=0.05, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_HDAC_Stress, TERM2NAME = pathways_KEGG_terms)


ORA_HDAC_KEGG_up@organism="bnapus"
ORA_HDAC_KEGG_up@ontology="KEGG"

ORA_Stress_KEGG_up@organism="bnapus"
ORA_Stress_KEGG_up@ontology="KEGG"

ORA_HDAC_Stress_KEGG_up@organism="bnapus"
ORA_HDAC_Stress_KEGG_up@ontology="KEGG"



HDAC_KEGG_UP <- importCP(ORA_HDAC_KEGG_up, type = "other")
write_csv(HDAC_KEGG_UP, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\ORA_PE_SAHA_UP_KEGG_Results.csv")

Stress_KEGG_UP <- importCP(ORA_Stress_KEGG_up, type = "other")
write_csv(Stress_KEGG_UP, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\ORA_HeatStress_UP_KEGG_Results.csv")

HDAC_Stress_KEGG_UP <- importCP(ORA_HDAC_Stress_KEGG_up, type = "other")
write_csv(HDAC_Stress_KEGG_UP, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\ORA_HeatStress_SAHA_UP_KEGG_Results.csv")


################# DOWN REGULATED GENES


bckgr_geneset_HDAC <- merge(background_DOWN_gl_HDAC, pathways_KEGG_gs, by="gene")
bckgr_geneset_Stress <- merge(background_DOWN_gl_Stress, pathways_KEGG_gs, by="gene")
bckgr_geneset_HDAC_Stress <- merge(background_DOWN_gl_HDAC_Stress, pathways_KEGG_gs, by="gene")

colnames(bckgr_geneset_HDAC)<-c("ID","bp")
bckgr_geneset_HDAC<-bckgr_geneset_HDAC[,c(2,1)]

colnames(bckgr_geneset_Stress)<-c("ID","bp")
bckgr_geneset_Stress<-bckgr_geneset_Stress[,c(2,1)]

colnames(bckgr_geneset_HDAC_Stress)<-c("ID","bp")
bckgr_geneset_HDAC_Stress<-bckgr_geneset_HDAC_Stress[,c(2,1)]

ORA_HDAC_KEGG_down <- clusterProfiler::enricher(gene = DEG_DOWN_HDAC, minGSSize=2, pvalueCutoff=0.05, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_HDAC, TERM2NAME = pathways_KEGG_terms)
ORA_Stress_KEGG_down <- clusterProfiler::enricher(gene = DEG_DOWN_Stress, minGSSize=2, pvalueCutoff=0.05, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_Stress, TERM2NAME = pathways_KEGG_terms)
ORA_HDAC_Stress_KEGG_down <- clusterProfiler::enricher(gene = DEG_DOWN_HDAC_Stress, minGSSize=2, pvalueCutoff=0.05, pAdjustMethod="BH", gson=NULL, TERM2GENE = bckgr_geneset_HDAC_Stress, TERM2NAME = pathways_KEGG_terms)


ORA_HDAC_KEGG_down@organism="bnapus"
ORA_HDAC_KEGG_down@ontology="KEGG"

ORA_Stress_KEGG_down@organism="bnapus"
ORA_Stress_KEGG_down@ontology="KEGG"

ORA_HDAC_Stress_KEGG_down@organism="bnapus"
ORA_HDAC_Stress_KEGG_down@ontology="KEGG"


HDAC_KEGG_DOWN <- importCP(ORA_HDAC_KEGG_down, type = "other")
write_csv(HDAC_KEGG_DOWN, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\ORA_PE_SAHA_DOWN_KEGG_Results.csv")

Stress_KEGG_DOWN <- importCP(ORA_Stress_KEGG_down, type = "other")
write_csv(Stress_KEGG_DOWN, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\ORA_HeatStress_DOWN_KEGG_Results.csv")

HDAC_Stress_KEGG_DOWN <- importCP(ORA_HDAC_Stress_KEGG_down, type = "other")
write_csv(HDAC_Stress_KEGG_DOWN, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\ORA_HeatStress_SAHA_DOWN_KEGG_Results.csv")

```

Post processing output and calculating FoldEnrich and RichFactor output
```{r}

# Calculate similarity matrix
simMatrix_HDACi <- calculateSimMatrix(HDAC_down$ID,
                                orgdb=org.Brassicanapus.eg.db,
                                ont="BP",
                                method="Wang")
# Reducing terms
scores <- setNames(-log10(HDAC_down$qvalue), HDAC_down$ID)
reducedTerms_HDAC <- reduceSimMatrix(simMatrix_HDACi,
                                scores,
                                threshold=0.7,
                                orgdb=org.Brassicanapus.eg.db)


##### STRESS

# Calculate similarity matrix
simMatrix_Stress <- calculateSimMatrix(Stress_down$ID,
                                orgdb=org.Brassicanapus.eg.db,
                                ont="BP",
                                method="Wang")
# Reducing terms
scores <- setNames(-log10(Stress_down$qvalue), Stress_down$ID)
reducedTerms_Stress <- reduceSimMatrix(simMatrix_Stress,
                                scores,
                                threshold=0.7,
                                orgdb=org.Brassicanapus.eg.db)
colnames(HDAC_down) <- c("BN_BP_ID", "Description", "GeneRatio", "BgRatio", "pvalue", "p.adjust", "qvalue", "geneID", "geneID_symbol", "Count", "FoldEnrich", "RichFactor")
colnames(Stress_down) <- c("BN_BP_ID", "Description", "GeneRatio", "BgRatio", "pvalue", "p.adjust", "qvalue", "geneID", "geneID_symbol", "Count", "FoldEnrich", "RichFactor")


write.table(reducedTerms_HDAC, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\reduced_HDAC_down.tsv", row.names=FALSE, quote=FALSE, sep="\t")
write.table(reducedTerms_Stress, "C:\\Users\\naata\\Documents\\MASTER\\TFM\\reduced_Stress_down.tsv", row.names=FALSE, quote=FALSE, sep="\t")

ORA_HDAC_ALL = ORA_HDAC_ALL[,]
ORA_Stress = ORA_Stress[,]

HDAC_Stress_KEGG_UP[!(HDAC_Stress_KEGG_UP$ID %in% Stress_KEGG_UP$ID),]$ID
```




```{r}
search_kegg_organism('bna', by='kegg_code')
```
```{r}
head(logFC)
map_file=read.csv("C:\\Users\\naata\\Documents\\MASTER\\TFM\\bna_EID.csv")
logFC_HDSAHAplot = merge(logFC_HDAC_Stress[,c(4,9)], map_file, by="external_gene_name")
```
```{r}
data(korg, package="pathview")
korg[which(korg[,3] == "bna"),]

```


```{r  pressure, echo=FALSE}

logFCv <- logFC_HDSAHAplot$log2FoldChange
data_kegg <- logFC_HDSAHAplot[,c(3,2)]
data_kegg=data_kegg[which(!duplicated(data_kegg[,1])),]
#names(data_kegg) <- logFC$external_gene_name 



data_final = t(data_final)
names(data_final)=data_final[1,]



pathview(gene.data = data_final,
         pathway.id = "bna03410", 
         species = "bna", # brassica napus KEGG accession 
         limit = list(gene=5, cpd=1),
         gene.idtype = "KEGG")

pathview(gene.data = data_final,
         pathway.id = "bna03410", 
         species = "bna", # brassica napus KEGG accession 
         limit = list(gene=5, cpd=1),
         gene.idtype = "KEGG")

pathview(gene.data = data_final,
         pathway.id = "bna03430", 
         species = "bna", # brassica napus KEGG accession 
         limit = list(gene=5, cpd=1),
         gene.idtype = "KEGG")

logFC_HDSAHAplot = merge(logFC_Stress[,c(3,9)], map_file, by="external_gene_name")

logFCv <- logFC_HDSAHAplot$log2FoldChange
data_kegg <- logFC_HDSAHAplot[,c(3,2)]
data_kegg=data_kegg[which(!duplicated(data_kegg[,1])),]
#names(data_kegg) <- logFC$external_gene_name 



data_final = t(data_final)
names(data_final)=data_final[1,]



pathview(gene.data = data_final,
         pathway.id = "bna03440", 
         species = "bna", # brassica napus KEGG accession 
         limit = list(gene=5, cpd=1),
         gene.idtype = "KEGG")

pathview(gene.data = data_final,
         pathway.id = "bna03008", 
         species = "bna", # brassica napus KEGG accession 
         limit = list(gene=5, cpd=1),
         gene.idtype = "KEGG")
```

