---
title: "Visualizations_Enrichment_ORA"
author: "Natalia García"
date: '2023-04-25'
output: html_document
---



```{r setup, include=FALSE}
#install.packages("stringi")
#library("biomartr")
library("multienrichjam")
library("clusterProfiler")
library("tidyverse")
library("enrichplot")
suppressPackageStartupMessages(library("org.At.tair.db"))
library("biomaRt")  # only use to remove cache bug
library("readxl")
```

# Load expression data
```{r}
logFC<-read_excel("C:\\Users\\naata\\MASTER\\Enrichment\\Filtered_DEG_LFC1_padj05_results_total.xlsx")
```

# Load and preprocess ShinyGO ORA results

(It is necessary to convert the tables to )

```{r cars}

# Reformat for clusterprofiler result intake

my_data <- read_excel("C:\\Users\\naata\\MASTER\\Enrichment\\ShinyGO_ORA_enrichment_HDAC.xlsx", sheet="enrichment_GO_BP")

ONT<-rep("BP",length(my_data[,1]))
my_data <- (cbind(my_data, ONT))

my_data$ID = my_data[,1]
rownames(my_data)= my_data[,1]

GeneRatio<-paste(my_data$nGenes, my_data[,4], sep="/")
BgRatio<-paste(my_data$nGenes,65732, sep="/")


my_data <- (cbind(my_data, GeneRatio))
my_data <- (cbind(my_data, BgRatio))

BP_GO_HDAC<-my_data[,c(9,10,6,11,12,2,4)]

BP_GO_HDAC$`p.adjust` <- BP_GO_HDAC[,6]
BP_GO_HDAC$`qvalue` <- BP_GO_HDAC[,6]
names(BP_GO_HDAC)<-c("ONTOLOGY", "ID", "Description", "GeneRatio", "BgRatio", "pvalue","p.adjust","qvalue", "FoldEnrich")

BP_GO_HDAC$intersections<-str_replace_all(my_data$Genes, "  ", ",")

BP_GO_HDAC$pvalue <- as.numeric(BP_GO_HDAC$pvalue)
BP_GO_HDAC$`p.adjust` <- as.numeric(BP_GO_HDAC$`p.adjust`)
BP_GO_HDAC$qvalue <- as.numeric(BP_GO_HDAC$qvalue)

```

Transform into enrich object

```{r}
enr <- enrichDF2enrichResult(enrichDF = BP_GO_HDAC, keyColname = "ID", geneColname = "intersections", pvalueColname = "pvalue", descriptionColname = "Description", pvalueCutoff = 0.05)
```


# Do the same for Stress dataset

```{r cars}

# Reformat for clusterprofiler result intake

my_data <- read_excel("C:\\Users\\naata\\MASTER\\Enrichment\\ShinyGO_ORA_Stress_Results.xlsx", sheet="enrichment_GO_BP_Stress")

ONT<-rep("BP",length(my_data[,1]))
my_data <- (cbind(my_data, ONT))

my_data$ID = my_data[,1]
rownames(my_data)= my_data[,1]

GeneRatio<-paste(my_data$nGenes, my_data[,4], sep="/")
BgRatio<-paste(my_data$nGenes,65732, sep="/")


my_data <- (cbind(my_data, GeneRatio))
my_data <- (cbind(my_data, BgRatio))

BP_GO_Stress<-my_data[,c(9,10,6,11,12,2,4)]

BP_GO_Stress$`p.adjust` <- BP_GO_Stress[,6]
BP_GO_Stress$`qvalue` <- BP_GO_Stress[,6]
names(BP_GO_Stress)<-c("ONTOLOGY", "ID", "Description", "GeneRatio", "BgRatio", "pvalue","p.adjust","qvalue", "FoldEnrich")

BP_GO_Stress$intersections<-str_replace_all(my_data$Genes, "  ", ",")

BP_GO_Stress$pvalue <- as.numeric(BP_GO_Stress$pvalue)
BP_GO_Stress$`p.adjust` <- as.numeric(BP_GO_Stress$`p.adjust`)
BP_GO_Stress$qvalue <- as.numeric(BP_GO_Stress$qvalue)

```

Transform into enrich object

```{r}
enr_S <- enrichDF2enrichResult(enrichDF = BP_GO_Stress, keyColname = "ID", geneColname = "intersections", pvalueColname = "pvalue", descriptionColname = "Description", pvalueCutoff = 0.05)
```
```{r}
ncol(enr)
```

```{r}
multienr <- data.frame(rows=c(nrow(enr), nrow(enr_S)),cols=c(ncol(enr), ncol(enr_S)), colnames<- c("HDAC", "Stress"))
```






























```{r}
ora_analysis_bp <- pairwise_termsim(enr, method = "JC")
emapplot(ora_analysis_bp, showCategory = 10, color = "qvalue", node_label_size = 0.0001)
```
```{r}
BP_GO_HDAC$geneID <- str_replace_all(BP_GO_HDAC[,9], "," ,"/")
BP_new<-BP_GO_HDAC[,c(2,3,4,5,6,7,8,10)]
names(BP_new)<-c("Bn_BP_ID","Description","GeneRatio","BgRatio","pvalue","p.adjust","qvalue","geneID")


library(igraph)
library(ggraph)
library("genekitr")
BP_new_discard<-BP_new[which(BP_new$Bn_BP_ID!="GO:0031497"),]
BP_new_discard<-BP_new[which(BP_new$pvalue<0.05),][1:10,]

svg("C:\\Users\\naata\\Desktop\\t10_GO_BP_terms_ORA.svg")
genekitr::plotEnrich(BP_new, plot_type = "gomap", term_metric = "GeneRatio", stats_metric="pvalue", up_color = '#a32a31',down_color = '#3665a6', ont="BP", org="bna")
dev.off()
```

```{r}
# Visualizations of gene-concept network
# library("AnnotationForge")
# library("AnnotationDbi")
# library("AnnotationHub")
# 
# # Making annotation package
# ah <- AnnotationHub()
# query(ah, c( "OrgDb", "Brassica napus"))
# org.Brassicanapus.eg.db <- ah[["AH107389"]]
# file.copy(AnnotationHub::cache(ah[["AH107389"]], "./org.Brassicanapus.eg.db"))
# 
# 
# makeOrgPackage(gene_info=fSym, chromosome=fChr, go=fGO,
#                version="0.1",
#                maintainer="Some One <so@someplace.org>",
#                author="Some One <so@someplace.org>",
#                outputDir = ".",
#                tax_id="59729",
#                genus="Brassica",
#                species="naous",
#                goTable="go",
#                verbose=T)


library(enrichplot)
## convert gene ID to Symbol
# bn_map <- setReadable(enr, org.Brassicanapus.eg.db, 'ENTREZID')
p1 <- cnetplot(enr)
## categorySize can be scaled by 'pvalue' or 'geneNum'
p2 <- cnetplot(enr, categorySize="pvalue", circular=TRUE, colorEdge = TRUE)

# p3 <- cnetplot(enr, foldChange=geneList, circular = TRUE, colorEdge = TRUE) 
# cowplot::plot_grid(p1, p2, p3, ncol=3, labels=LETTERS[1:3], rel_widths=c(.8, .8, 1.2))

```




```{r}
# tree plot
edox2 <- pairwise_termsim(enr)
p1 <- treeplot(edox2)
p1
#p2 <- treeplot(edox2, hclust_method = "average")
#aplot::plot_list(p1, p2, tag_levels='A')
```
```{r}
#xx <- compareCluster(gcSample, fun="enrichGO", organism="hsa", pvalueCutoff=0.05)
#emapplot(edox2, pie="count", #cex_category=0.5, layout="kk")
```



DNA methylation cytosine
#"GO:0031497" --> GO:0006325
#"GO:0006333" --> GO:0006325
#"GO:0006323" --> GO:0051276
#"GO:0070919" --> GO:0030422
#"GO:0016246" --> GO:0035194
#"GO:0031050" --> GO:0031050

**GOMAP**

```{r}
#BP_GO_HDAC$geneID <- str_replace_all(BP_GO_HDAC[,9], "," ,"/")
#BP_new<-BP_GO_HDAC[,c(2,3,4,5,6,7,8,10)]
#names(BP_new)<-c("Bn_BP_ID","Description","GeneRatio","BgRatio","pvalue","p.adjust","qvalue","geneID")
# BP_new_epig<-subset(BP_new, (Bn_BP_ID %in% c("GO:0051276","GO:0031507","GO:0045814","GO:0031048","GO:0070828","GO:0006325","GO:0040029","GO:0071103","GO:0044728","GO:0006305","GO:0006306","GO:0065004","GO:0006338","GO:0032776","GO:0071824","GO:0030261","GO:0031047","GO:0031048","GO:0035194","GO:0030422","GO:0070918","GO:0031167","GO:0044815","GO:0032993","GO:0000785","GO:0000123","GO:0032777","GO:0004402")))

#library(igraph)
#library(ggraph)
#library("genekitr")


BP_new_dna_met_cytosine<-subset(BP_new, (Bn_BP_ID %in% c("GO:0032776","GO:0006306","GO:0006305","GO:00060304","GO:0043414","GO:0006259","GO:0090304","GO:0006139","GO:0043412","GO:0006725","GO:0046483","GO:0044238","GO:0034641","GO:1901360","GO:0043170","GO:0032259","GO:0071704","GO:0006807","GO:0044237","GO:0008152","GO:0009987","GO:0008150","GO:0044815")))


svg("C:\\Users\\naata\\Desktop\\GO_BP_terms_ORA_dnametC.svg")
genekitr::plotEnrich(BP_new_dna_met_cytosine, plot_type = "gomap", term_metric = "GeneRatio", stats_metric="pvalue", up_color = '#a32a31',down_color = '#3665a6', ont="BP", org="bna") + coord_cartesian( expand=TRUE)
dev.off()
```
**GENE NET**
```{r}

# Import external gene name - log2C mappings
geneFC<-logFC[which(!duplicated(logFC$external_gene_name)),3]
listfc<-as.list(t(geneFC))
names(listfc) = as.list(t(logFC[which(!duplicated(logFC$external_gene_name)),8]))


enr_2 <- enrichDF2enrichResult(enrichDF = BP_new_dna_met_cytosine, keyColname = "Bn_BP_ID", geneColname = "geneID", pvalueColname = "pvalue", descriptionColname = "Description", pvalueCutoff = 0.05)


svg("C:\\Users\\naata\\Desktop\\genenet.svg")
cnetplot(enr_2, categorySize="GeneRatio", foldChange = listfc, colorEdge = TRUE, node_label = 'gene', cex_gene=3)

```


ncRNA meiated post transcriptional gene silencing

```{r}


BP_new_ncRNAgs<-subset(BP_new, (Bn_BP_ID %in% c("GO:0035194","GO:0009892","GO:0060255","GO:0048519","GO:0043170","GO:0019222","GO:0071704","GO:0008152","GO:0065007","GO:0008150")))


svg("C:\\Users\\naata\\Desktop\\GO_BP_terms_ORA_ncRNAgs.svg")
genekitr::plotEnrich(BP_new_ncRNAgs, plot_type = "gomap", term_metric = "GeneRatio", stats_metric="pvalue", up_color = '#a32a31',down_color = '#3665a6', ont="BP", org="bna") + coord_cartesian( expand=TRUE)
dev.off()
```
siRNA silencing

```{r}


BP_new_siRNA<-subset(BP_new, (Bn_BP_ID %in% c("GO:0008150","GO:0009987","GO:000812","GO:0065007","GO:0044237","GO:0006807","GO:0071704","GO:0050789","GO:00046483","GO:0006725", "GO:0034641", "GO:1901360", "GO:0044238", "GO:0043170", "GO:0019222", "GO:0048519", "GO:0006139", "GO:0010467", "GO:0060255", "GO:0009892", "GO:0090304", "GO:0010468", "GO:0010605", "GO:0016070", "GO:0010629", "GO:0034660", "GO:0006396", "GO:0034470", "GO:0031047", "GO:0070918", "GO:0030422" )))


svg("C:\\Users\\naata\\Desktop\\GO_BP_terms_ORA_ncRNAgs.svg")
genekitr::plotEnrich(BP_new_siRNA, plot_type = "gomap", term_metric = "GeneRatio", stats_metric="pvalue", up_color = '#a32a31',down_color = '#3665a6', ont="BP", org="bna") + coord_cartesian( expand=TRUE)
dev.off()
```

```{r}

enr_3 <- enrichDF2enrichResult(enrichDF = BP_new_siRNA, keyColname = "Bn_BP_ID", geneColname = "geneID", pvalueColname = "pvalue", descriptionColname = "Description", pvalueCutoff = 0.05)


svg("C:\\Users\\naata\\Desktop\\overlap_genenet_siRNA_HDAC.svg")
cnetplot(enr_3, categorySize="GeneRatio", foldChange = listfc, colorEdge = TRUE, node_label = 'none', cex_gene=3, showCategory = 10)
dev.off()
```


```{r}
B_selected_epi_specific = subset(BP_new, (Bn_BP_ID %in% c( "GO:0004402", "GO:0031048", "GO:0045814", "GO:0030422" , "GO:0031507", "GO:0040029", "GO:0000123", "GO:0032777" )))

enr_4 <- enrichDF2enrichResult(enrichDF = B_selected_epi_specific, keyColname = "Bn_BP_ID", geneColname = "geneID", pvalueColname = "pvalue", descriptionColname = "Description", pvalueCutoff = 0.05)


svg("C:\\Users\\naata\\Desktop\\selected_epi_genenet.svg")
cnetplot(enr_4, categorySize="GeneRatio", foldChange = listfc, colorEdge = TRUE, node_label = 'gene', cex_gene=3, showCategory = 10)
dev.off()
#genekitr::plotEnrich(B_selected_epi_specific, plot_type = "gomap", term_metric = "GeneRatio", stats_metric="pvalue", up_color = '#a32a31',down_color = '#3665a6', ont="BP", org="bna") + coord_cartesian( expand=TRUE)
```

```{r}
B_selected_aux_specific = subset(BP_new, (Bn_BP_ID %in% c( "GO:0006586", "GO:0042430", "GO:0010928", "GO:0010315" , "GO:0010252", "GO:0042445", "GO:0009926", "GO:0009733","GO:0060918", "GO:0009914", "GO:0010817", "GO:0071365", "GO:0009734", "GO:0071495", "GO:0032870", "GO:0009755", "GO:0009719", "GO:0009725" )))

enr_5 <- enrichDF2enrichResult(enrichDF = B_selected_aux_specific, keyColname = "Bn_BP_ID", geneColname = "geneID", pvalueColname = "pvalue", descriptionColname = "Description", pvalueCutoff = 0.05)


svg("C:\\Users\\naata\\Desktop\\selected_aux_genenet.svg")
cnetplot(enr_5, categorySize="GeneRatio", foldChange = listfc, colorEdge = TRUE, node_label = 'none', cex_gene=3, showCategory = 10)
dev.off()
#genekitr::plotEnrich(B_selected_epi_specific, plot_type = "gomap", term_metric = "GeneRatio", stats_metric="pvalue", up_color = '#a32a31',down_color = '#3665a6', ont="BP", org="bna") + coord_cartesian( expand=TRUE)
```
```{r}
svg("C:\\Users\\naata\\Desktop\\GO_BP_terms_ORA_aux.svg")
genekitr::plotEnrich(B_selected_aux_specific, plot_type = "gomap", term_metric = "GeneRatio", stats_metric="pvalue", up_color = '#a32a31',down_color = '#3665a6', ont="BP", org="bna") + coord_cartesian( expand=TRUE)
dev.off()
```

```{r}
B_selected_cgrowth_specific = subset(BP_new, (Bn_BP_ID %in% c( "GO:0008283", "GO:0016049", "GO:0001558", "GO:0009826" , "GO:0042814", "GO:0042814", "GO:0060560", "GO:0051510","GO:0051301", "GO:0008356", "GO:0009786", "GO:0032502", "GO:0007275", "GO:0048731", "GO:0003006", "GO:0009791", "GO:0050793", "GO:0048582", "GO:0060560")))

enr_6 <- enrichDF2enrichResult(enrichDF = B_selected_cgrowth_specific, keyColname = "Bn_BP_ID", geneColname = "geneID", pvalueColname = "pvalue", descriptionColname = "Description", pvalueCutoff = 0.05)


svg("C:\\Users\\naata\\Desktop\\selected_cg_genenet.svg")
cnetplot(enr_6, categorySize="GeneRatio", foldChange = listfc, colorEdge = TRUE, node_label = 'none', cex_gene=3, showCategory = 10)
dev.off()
#genekitr::plotEnrich(B_selected_epi_specific, plot_type = "gomap", term_metric = "GeneRatio", stats_metric="pvalue", up_color = '#a32a31',down_color = '#3665a6', ont="BP", org="bna") + coord_cartesian( expand=TRUE)

svg("C:\\Users\\naata\\Desktop\\GO_BP_terms_ORA_cprolif_growth.svg")
genekitr::plotEnrich(B_selected_cgrowth_specific, plot_type = "gomap", term_metric = "GeneRatio", stats_metric="pvalue", up_color = '#a32a31',down_color = '#3665a6', ont="BP", org="bna") + coord_cartesian( expand=TRUE)
dev.off()
```


```{r}
#library(enrichR)
library(patchwork)



# Auxin reg






#p3 <- genekitr::plotEnrich(enr, plot_type = "geneheat", show_gene = show_gene_aux, fold_change = listfc)

#
p3 <- heatplot(ego, show_gene = show_gene_aux, foldChange = listfc)

```




```{r}
plotEnrich(enr,
           plot_type = 'dot',
           scale_ratio = 2, # dot size
           main_text_size = 10,
           legend_text_size =8,
           n_term = 6) # show terms
```


---
# KEGG


#Script to visualize overrepresented KEGG pathways

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading libraries

```{r}
library("pathview")
library("readxl")
library("clusterprofiler")
library("Rgraphviz")
library("graph")
library("KEGGgraph")
```


## Loading data
```{r}
keggpath_DEG<-read_excel("C:\\Users\\naata\\MASTER\\Enrichment\\ShinyGO_ORA_enrichment_HDAC.xlsx", sheet = "enrichment_KEGG")
```
# Plotting first 


```{r}
search_kegg_organism('bna', by='kegg_code')
```
```{r}
head(logFC)
```
```{r}
data(korg, package="pathview")
korg[which(korg[,3] == "bna"),]
head(logFC)
```


```{r  pressure, echo=FALSE}

logFCv <- logFC$log2FoldChange
data_kegg <- logFC[,c(8,3)]
data_kegg=data_kegg[which(!duplicated(data_kegg[,1])),]
#names(data_kegg) <- logFC$external_gene_name 

# Annotating DEGs
# we want the correspondence of TAIR/Ensembl symbols with NCBI Entrez gene ids

attributes_to_retrieve = c("external_gene_name","entrezgene_id")
# see a list of "marts" available at host "plants.ensembl.org"
listMarts <-listMarts(host="https://plants.ensembl.org")

# create an object 
mart <- useMart(biomart="plants_mart", host="https://plants.ensembl.org")

# see a list of data sets within the mart

listdatasets <- listDatasets(mart)
(listDatasets(mart))[grep("bnapus",listDatasets(mart)[,1]),]

# create an object for the plants_mart-Ensembl 

bnapus_mart <- useMart(biomart = "plants_mart", host = "https://plants.ensembl.org", dataset = "bnapus_eg_gene"	)


# get gene names and transcript lengths when they exist
ann <- getBM(filter="",value= "",attributes=c("external_gene_name","entrezgene_id"), mart=bnapus_mart, useCache = FALSE)


# for compatibility with enrichGO universe
# genes in the universe need to be characters and not integers (Entrez gene id)
ann$entrezgene_id = as.character(ann$entrezgene_id)

data_final<- merge(data_kegg, ann, by="external_gene_name", all.x=FALSE, all.y=TRUE)

data_final = t(data_final)
names(data_final)=data_final[1,]



pathview(gene.data = data_final,
         pathway.id = "bna04075", 
         species = "bna", # brassica napus KEGG accession 
         limit = list(gene=5, cpd=1),
         gene.idtype = "KEGG")
```



## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
